---
title: "Multiomics TB biomarker discovery project"
author: "Zaynab Mousavian"
date: "September 1, 2026"
output:
  html_document: default
  pdf_document: default
subtitle: Making Tables
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Load R packages

```{r}
library(limma)
library(ggplot2)
library(tidyverse)
library(ggrepel)
library(ComplexHeatmap)
library(circlize)
library(dplyr)
library(grid)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(caret)
library(randomForestSRC)
library(pROC)
library(glmnet)
library(Boruta)
library(doParallel)
library(patchwork)
library(corto)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(caret)
library(ggforce)   
library(ggh4x)
library(gridGraphics)
library(cowplot)
library(Hmisc)
library(readxl)
library(caret)
library(pROC)
library(doParallel)
library(foreach)
library(dplyr)
```



# Make Table 1
## load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult[match(row.names(CytokineData_Baseline_Adult_imputed),ClinicalData_Baseline_Adult$ID),]
dim(CytokineData_Baseline_Adult)
```

## Age

```{r}
ClinicalData_Baseline_Adult <-ClinicalData_Baseline_Adult %>%
  mutate(SA = if_else(Study == "SA", "SA", "noSA"))

age_summary <- ClinicalData_Baseline_Adult %>%
  group_by(TB, SA) %>%
  summarise(
    med = median(AGE, na.rm = TRUE),
    Q1  = quantile(AGE, 0.25, na.rm = TRUE),
    Q3  = quantile(AGE, 0.75, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`Median (IQR)` = paste0(med, " (", Q1, "–", Q3, ")"))

# Create total by collapsing over site
age_total <- ClinicalData_Baseline_Adult %>%
  group_by(TB) %>%      # only TB groups, ignore site
  summarise(
    med = median(AGE, na.rm = TRUE),
    Q1  = quantile(AGE, 0.25, na.rm = TRUE),
    Q3  = quantile(AGE, 0.75, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    site = "Total",
    `Median (IQR)` = paste0(med, " (", Q1, "–", Q3, ")")
  )
```

## Sex
```{r}
female_counts <- ClinicalData_Baseline_Adult %>%
  filter(Sex == "F") %>%
  group_by(TB, SA) %>%
  summarise(N = n(), .groups = "drop")

tb_totals <- ClinicalData_Baseline_Adult %>%
  filter(Sex == "F") %>%
  group_by(TB) %>%
  summarise(Total = n(), .groups = "drop")

table_data <- female_counts %>%
  left_join(tb_totals, by = "TB") %>%
  mutate(Percent = paste0(N, " (", round(N / Total * 100, 1), "%)"))

final_table <- data.frame(
  TB_SA = table_data$Percent[table_data$TB=="Yes" & table_data$SA=="SA"],
  TB_noSA = table_data$Percent[table_data$TB=="Yes" & table_data$SA=="noSA"],
  TB_Total = table_data$Percent[table_data$TB=="Yes" & table_data$SA %in% c("SA","noSA")] %>% sum(., na.rm = TRUE),
  noTB_SA = table_data$Percent[table_data$TB=="No" & table_data$SA=="SA"],
  noTB_noSA = table_data$Percent[table_data$TB=="No" & table_data$SA=="noSA"],
  noTB_Total = table_data$Percent[table_data$TB=="No" & table_data$SA %in% c("SA","noSA")] %>% sum(., na.rm = TRUE)
)
```

## HIV

```{r}
HIV_counts <- ClinicalData_Baseline_Adult %>%
  filter(HIV=="Positive") %>%
  group_by(TB, SA) %>%
  summarise(N = n(), .groups = "drop")

tb_totals <- ClinicalData_Baseline_Adult %>%
  filter(HIV=="Positive") %>%
  group_by(TB) %>%
  summarise(Total = n(), .groups = "drop")

table_data <- female_counts %>%
  left_join(tb_totals, by = "TB") %>%
  mutate(Percent = paste0(N, " (", round(N / Total * 100, 1), "%)"))

final_table <- data.frame(
  TB_SA = table_data$Percent[table_data$TB=="Yes" & table_data$SA=="SA"],
  TB_noSA = table_data$Percent[table_data$TB=="Yes" & table_data$SA=="noSA"],
  TB_Total = table_data$Percent[table_data$TB=="Yes" & table_data$SA %in% c("SA","noSA")] %>% sum(., na.rm = TRUE),
  noTB_SA = table_data$Percent[table_data$TB=="No" & table_data$SA=="SA"],
  noTB_noSA = table_data$Percent[table_data$TB=="No" & table_data$SA=="noSA"],
  noTB_Total = table_data$Percent[table_data$TB=="No" & table_data$SA %in% c("SA","noSA")] %>% sum(., na.rm = TRUE)
)
```

## CD4 for HIV positive

```{r}
ClinicalData_Baseline_Adult %>%
  filter(HIV == "Positive") %>%
  group_by(TB, SA) %>%
  summarise(
    CD4 = paste0(
      median(CD4, na.rm = TRUE),
      " (",
      quantile(CD4, 0.25, na.rm = TRUE),
      "–",
      quantile(CD4, 0.75, na.rm = TRUE),
      ")"
    )
  )

ClinicalData_Baseline_Adult %>%
  filter(HIV == "Positive") %>%
  group_by(TB) %>%
  summarise(
    CD4 = paste0(
      median(CD4, na.rm = TRUE),
      " (",
      quantile(CD4, 0.25, na.rm = TRUE),
      "–",
      quantile(CD4, 0.75, na.rm = TRUE),
      ")"
    )
  )

ClinicalData_Baseline_Adult %>%
  filter(HIV == "Positive") %>%
  group_by(TB) %>%
  summarise(
    n_missing_CD4 = sum(is.na(CD4))
  )
```

## HIV Viral load for HIV positive

```{r}
ClinicalData_Baseline_Adult %>% 
  filter(HIV == "Positive") %>%
  mutate(
    VL_numeric = case_when(
      as.character(HIV.Viral.Load) == "LDL" ~ 20,
      TRUE ~ as.numeric(as.character(HIV.Viral.Load))
    )
  )%>%
  group_by(TB, SA) %>%
  summarise(
    VL = paste0(
      median(VL_numeric, na.rm = TRUE),
      " (",
      quantile(VL_numeric, 0.25, na.rm = TRUE),
      "–",
      quantile(VL_numeric, 0.75, na.rm = TRUE),
      ")"
    )
  )

ClinicalData_Baseline_Adult %>% 
  filter(HIV == "Positive") %>%
  mutate(
    VL_numeric = case_when(
      as.character(HIV.Viral.Load) == "LDL" ~ 20,
      TRUE ~ as.numeric(as.character(HIV.Viral.Load))
    )
  )%>%
  group_by(TB) %>%
  summarise(
    VL = paste0(
      median(VL_numeric, na.rm = TRUE),
      " (",
      quantile(VL_numeric, 0.25, na.rm = TRUE),
      "–",
      quantile(VL_numeric, 0.75, na.rm = TRUE),
      ")"
    )
  )
ClinicalData_Baseline_Adult %>%
  filter(HIV == "Positive") %>%
  group_by(TB) %>%
  summarise(
    n_missing_VL = sum(is.na(HIV.Viral.Load))
  )
```
## CXray

```{r}
HIV_counts <- ClinicalData_Baseline_Adult %>%
  filter(cxray=="CAVITATION") %>%
  group_by(TB, SA) %>%
  summarise(N = n(), .groups = "drop")

tb_totals <- ClinicalData_Baseline_Adult %>%
  filter(cxray=="CAVITATION") %>%
  group_by(TB) %>%
  summarise(Total = n(), .groups = "drop")


HIV_counts <- ClinicalData_Baseline_Adult %>%
  filter(Speciation.Results=="DETECTED") %>%
  group_by(TB, SA) %>%
  summarise(N = n(), .groups = "drop")

tb_totals <- ClinicalData_Baseline_Adult %>%
  filter(Speciation.Results=="DETECTED") %>%
  group_by(TB) %>%
  summarise(Total = n(), .groups = "drop")

HIV_counts <- ClinicalData_Baseline_Adult %>%
  filter(Sputum.Results=="Positive") %>%
  group_by(TB, SA) %>%
  summarise(N = n(), .groups = "drop")

tb_totals <- ClinicalData_Baseline_Adult %>%
  filter(Sputum.Results=="Positive") %>%
  group_by(TB) %>%
  summarise(Total = n(), .groups = "drop")

HIV_counts <- ClinicalData_Baseline_Adult %>%
  filter(Sputum.smear==2) %>%
  group_by(TB, SA) %>%
  summarise(N = n(), .groups = "drop")

tb_totals <- ClinicalData_Baseline_Adult %>%
  filter(Sputum.smear==2) %>%
  group_by(TB) %>%
  summarise(Total = n(), .groups = "drop")
```

## Symptoms 

```{r}
diagnosis <- read_excel("../../Data/ORD samples with diagnoses .xlsx")
SA_symptoms <- read_excel("../../Data/2025-03-03_PotPrev_Emory_Data.xlsx")
GWAS_symptoms <- read_excel("../../Data/TBGWAS_ControlSamples_30Nov2023.xlsx")
SA_symptoms$study_id <- paste0("SA",SA_symptoms$study_id)
GWAS_symptoms$`subjid without dash` <- paste0("TBG",GWAS_symptoms$`subjid without dash`)
samples <- ClinicalData_Baseline_Adult%>%filter(TB=="No")%>%select(ID)%>%unlist()

df1 <- SA_symptoms %>%
  filter(study_id %in% samples)

symptoms1 <- c("Cough","Fever...24", "Loss of weight","Night sweats")

sa_summary <- df1 %>%
  select(all_of(symptoms1)) %>%
  pivot_longer(everything(),
               names_to = "Symptom",
               values_to = "Response") %>%
  filter(!is.na(Response)) %>%
  group_by(Symptom) %>%
  summarise(
    n1 = sum(Response == "Yes"),
    N1 = n(),
    .groups = "drop"
  )
sa_summary

df2 <- GWAS_symptoms %>%
  filter(`subjid without dash` %in% samples)

symptoms2 <- c("a1_cough_c_bl","a1_fever_c_bl","a1_weightloss_c_bl",
               "a1_nightsweats_c_bl")

gwas_summary <- df2 %>%
  select(all_of(symptoms2)) %>%
  pivot_longer(everything(),
               names_to = "Symptom",
               values_to = "Response") %>%
  filter(!is.na(Response)) %>%
  group_by(Symptom) %>%
  summarise(
    n2 = sum(Response == "1"),
    N2 = n(),
    .groups = "drop"
  )
gwas_summary

gwas_summary$Symptom <- c("Cough","Fever...24",
                          "Loss of weight","Night sweats")

### Total
final_table <- sa_summary %>%
  inner_join(gwas_summary, by = "Symptom") %>%
  mutate(
    n_total = n1 + n2,
    N_total = N1 + N2,
    percent = round(n_total / N_total * 100, 1),
    `n (%)` = paste0(n_total, " (", percent, "%)")
  ) %>%
  select(Symptom, `n (%)`)

final_table

samples <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes")%>%select(ID)%>%unlist()

df1 <- SA_symptoms %>%
  filter(study_id %in% samples)

symptoms1 <- c("Cough","Fever...24", "Loss of weight","Night sweats")

sa_summary <- df1 %>%
  select(all_of(symptoms1)) %>%
  pivot_longer(everything(),
               names_to = "Symptom",
               values_to = "Response") %>%
  filter(!is.na(Response)) %>%
  group_by(Symptom) %>%
  summarise(
    n1 = sum(Response == "Yes"),
    N1 = n(),
    .groups = "drop"
  )
sa_summary
```

# Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()
```

# Make train and test data  


```{r message=FALSE, warning=FALSE}
ClinicalData_Baseline_Adult$Group <- cbind(paste0(ClinicalData_Baseline_Adult$TB,",",ClinicalData_Baseline_Adult$HIV))

set.seed(123) 
test_percentage <- 0.3 

test_indices <- createDataPartition(ClinicalData_Baseline_Adult$Group, p = test_percentage, list = FALSE)

sampleData_train <- ClinicalData_Baseline_Adult[-test_indices, ]
sampleData_test <- ClinicalData_Baseline_Adult[test_indices, ]


removed_samles_train <- which(!sampleData_train$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_train <- sampleData_train[-removed_samles_train,]
removed_samles_test <- which(!sampleData_test$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_test <- sampleData_test[-removed_samles_test,]

table(sampleData_train$TB,sampleData_train$HIV)
table(sampleData_test$TB,sampleData_test$HIV)

metaboData.train <- MetaboData_Baseline_Adult[sampleData_train$ID,]
cytokineData.train <- CytokineData_Baseline_Adult_imputed[sampleData_train$ID,]
sampleData.train <- sampleData_train
metaboData.test <- MetaboData_Baseline_Adult[sampleData_test$ID,]
cytokineData.test <- CytokineData_Baseline_Adult_imputed[sampleData_test$ID,]
sampleData.test <- sampleData_test


metaboData.train <- as.data.frame(metaboData.train)
cytokineData.train <- as.data.frame(cytokineData.train)

sampleData_test_HIVpos <- sampleData_test%>%
  filter(HIV=="Positive")
sampleData_test_HIVneg <- sampleData_test%>%
  filter(HIV=="Negative")
sampleData_test_noTBSA <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "SA"))
sampleData_test_noTBGWAS <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "TBGWAS"))

metaboData.test.HIVpos <- metaboData.test[sampleData_test_HIVpos$ID,]
cytokineData.test.HIVpos <- cytokineData.test[sampleData_test_HIVpos$ID,]
sampleData.test.HIVpos <- sampleData_test_HIVpos

metaboData.test.HIVneg <- metaboData.test[sampleData_test_HIVneg$ID,]
cytokineData.test.HIVneg <- cytokineData.test[sampleData_test_HIVneg$ID,]
sampleData.test.HIVneg <- sampleData_test_HIVneg

metaboData.test.noTBSA <- metaboData.test[sampleData_test_noTBSA$ID,]
cytokineData.test.noTBSA <- cytokineData.test[sampleData_test_noTBSA$ID,]
sampleData.test.noTBSA <- sampleData_test_noTBSA

metaboData.test.noTBGWAS <- metaboData.test[sampleData_test_noTBGWAS$ID,]
cytokineData.test.noTBGWAS <- cytokineData.test[sampleData_test_noTBGWAS$ID,]
sampleData.test.noTBGWAS <- sampleData_test_noTBGWAS
```

# Make a combined data including metabolites and cytokine for feature selection 

```{r}
combinedData.train <- cbind(metaboData.train,cytokineData.train)
combinedData.test <- cbind(metaboData.test,cytokineData.test)
combinedData.train$Group= as.factor(sampleData.train$TB)
combinedData.train$Group <- factor(combinedData.train$Group, levels =c("Yes","No"))
```

# Make Supplementary Table 1 

## Run all seuqntial combinations out of 13 markers selected by randomForestSRC

```{r}
library(caret)
library(pROC)
library(dplyr)
library(doParallel)

# ============================================================
# 13 sequential signatures
# ============================================================

features <- c(
  "IFN.gamma", "IL.1beta", "IL.22", "METHIONINE", "OXOPROLINE",
  "IL.10", "IFN.beta", "PYRUVATE", "IP.10", "IL.18",
  "TNF.alpha", "IL.6", "C22-5-FA"
)

set.seed(123)

# ------------------------------------------------------------
# SVM cross-validation settings
# ------------------------------------------------------------

fitControl <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 10,
  classProbs = TRUE,
  summaryFunction = twoClassSummary
)

svmGrid <- expand.grid(
  sigma = seq(0.01, 1, length.out = 10),
  C = 2^seq(-5, 10, length.out = 10)
)

# ------------------------------------------------------------
# Start parallel processing
# ------------------------------------------------------------

cl <- makeCluster(max(1, detectCores() - 1))
registerDoParallel(cl)

# ------------------------------------------------------------
# Run all 13 signatures
# ------------------------------------------------------------

results <- list()

for (k in 1:length(features)) {
  
  cat("Running signature", k, "of", length(features), "\n")
  
  # Features included in this signature
  selected_features <- features[1:k]
  
  # Identify cytokines and metabolites
  selected_cytokines <- selected_features[
    selected_features %in% colnames(cytokineData.train)
  ]
  
  selected_metabolites <- selected_features[
    selected_features %in% colnames(metaboData.train)
  ]
  
  # ----------------------------------------------------------
  # TRAINING DATA
  # ----------------------------------------------------------
  
  data_train <- cbind(
    cytokineData.train[, selected_cytokines, drop = FALSE],
    metaboData.train[, selected_metabolites, drop = FALSE]
  )
  
  data_train <- as.data.frame(data_train)
  
  data_train$Class <- factor(
    gsub(" ", "", sampleData.train$TB),
    levels = c("Yes", "No")
  )
  
  # ----------------------------------------------------------
  # TEST DATA
  # ----------------------------------------------------------
  
  data_test <- cbind(
    cytokineData.test[, selected_cytokines, drop = FALSE],
    metaboData.test[, selected_metabolites, drop = FALSE]
  )
  
  data_test <- as.data.frame(data_test)
  
  data_test$Class <- factor(
    gsub(" ", "", sampleData.test$TB),
    levels = c("Yes", "No")
  )
  
  # ----------------------------------------------------------
  # TRAIN SVM
  # ----------------------------------------------------------
  
  set.seed(123)
  
  svmTune <- train(
    Class ~ .,
    data = data_train,
    method = "svmRadial",
    preProc = c("center", "scale"),
    metric = "ROC",
    tuneGrid = svmGrid,
    trControl = fitControl
  )
  
  # ----------------------------------------------------------
  # PREDICT ON INDEPENDENT TEST SET
  # ----------------------------------------------------------
  
  prob_test <- predict(
    svmTune,
    newdata = data_test,
    type = "prob"
  )
  
  # Probability of TB
  ptb_prob <- prob_test[, "Yes"]
  
  # ----------------------------------------------------------
  # ROC / AUC
  # ----------------------------------------------------------
  
  roc_obj <- roc(
    response = data_test$Class,
    predictor = ptb_prob,
    levels = c("No", "Yes"),
    direction = "<",
    ci = TRUE,
    boot.n = 2000
  )
  
  auc_value <- as.numeric(auc(roc_obj))
  
  auc_ci <- as.numeric(
    ci.auc(
      roc_obj,
      conf.level = 0.95,
      boot.n = 2000
    )
  )
  
  # ----------------------------------------------------------
  # SENSITIVITY AT 70% SPECIFICITY
  # ----------------------------------------------------------
  
  ci_70 <- ci.se(
    roc_obj,
    specificities = 0.70,
    conf.level = 0.95,
    boot.n = 2000
  )
  
  sens_70 <- as.numeric(ci_70[1, 2])
  lower_70 <- as.numeric(ci_70[1, 1])
  upper_70 <- as.numeric(ci_70[1, 3])
  
  # ----------------------------------------------------------
  # SENSITIVITY AT 98% SPECIFICITY
  # ----------------------------------------------------------
  
  ci_98 <- ci.se(
    roc_obj,
    specificities = 0.98,
    conf.level = 0.95,
    boot.n = 2000
  )
  
  sens_98 <- as.numeric(ci_98[1, 2])
  lower_98 <- as.numeric(ci_98[1, 1])
  upper_98 <- as.numeric(ci_98[1, 3])
  
  # ----------------------------------------------------------
  # SAVE RESULTS
  # ----------------------------------------------------------
  
  results[[k]] <- data.frame(
    
    Signature = k,
    
    Markers = paste(
      selected_features,
      collapse = " + "
    ),
    
    AUC = auc_value,
    AUC_lower95 = auc_ci[1],
    AUC_upper95 = auc_ci[3],
    
    Sensitivity_70 = sens_70,
    Sensitivity_70_lower95 = lower_70,
    Sensitivity_70_upper95 = upper_70,
    
    Sensitivity_98 = sens_98,
    Sensitivity_98_lower95 = lower_98,
    Sensitivity_98_upper95 = upper_98
  )
}

# ------------------------------------------------------------
# Stop parallel processing
# ------------------------------------------------------------

stopCluster(cl)
registerDoSEQ()

# ------------------------------------------------------------
# Combine all 13 signatures
# ------------------------------------------------------------

results_13_signatures <- bind_rows(results)

# ------------------------------------------------------------
# Create manuscript-ready table
# ------------------------------------------------------------

results_table <- results_13_signatures %>%
  mutate(
    
    `AUC (95% CI)` = sprintf(
      "%.2f (%.2f–%.2f)",
      AUC,
      AUC_lower95,
      AUC_upper95
    ),
    
    `Sensitivity at 70% specificity (95% CI)` = sprintf(
      "%.2f%% (%.2f–%.2f%%)",
      Sensitivity_70 * 100,
      Sensitivity_70_lower95 * 100,
      Sensitivity_70_upper95 * 100
    ),
    
    `Sensitivity at 98% specificity (95% CI)` = sprintf(
      "%.2f%% (%.2f–%.2f%%)",
      Sensitivity_98 * 100,
      Sensitivity_98_lower95 * 100,
      Sensitivity_98_upper95 * 100
    )
  ) %>%
  select(
    Signature,
    Markers,
    `AUC (95% CI)`,
    `Sensitivity at 70% specificity (95% CI)`,
    `Sensitivity at 98% specificity (95% CI)`
  )

# ------------------------------------------------------------
# Display final table
# ------------------------------------------------------------

results_table
write.csv(results_table, "svm_combinations_results_sorted-13markers.csv", row.names = FALSE)
```

