---
title: "Multiomics TB biomarker discovery project"
author: "Zaynab Mousavian"
date: "March 1, 2026"
output:
  html_document: default
  pdf_document: default
subtitle: Making Figures
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Load R packages

```{r}
library(limma)
library(ggplot2)
library(tidyverse)
library(ggrepel)
library(ComplexHeatmap)
library(circlize)
library(dplyr)
library(grid)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(caret)
library(randomForestSRC)
library(pROC)
library(glmnet)
library(Boruta)
library(doParallel)
library(patchwork)
library(corto)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(caret)
library(ggforce)   
library(ggh4x)
library(gridGraphics)
library(cowplot)
library(Hmisc)
library(readxl)
```

# Functions {.tabset}
## Function for visaulizing volcano plot 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func <- function(inputData, xaxis, yaxis, uplab, downlab, name,
                             thr = 0.05, thr2 = 0, 
                             color_up = "red", color_down = "blue", color_nsig = "grey",
                             label_points = TRUE,
                             show_threshold_lines = TRUE,
                             xlim_range = NULL,
                             highlight_names=NULL) {
  
  inputData$sig <- ifelse(inputData[[yaxis]] > thr | abs(inputData[[xaxis]]) <= thr2, 
                          "Not Significant",
                          ifelse(inputData[[xaxis]] > thr2, "Up", "Down"))
  inputData$sig <- factor(inputData$sig,
                          levels = c("Up", "Down", "Not Significant"))
  
  p <- ggplot(inputData, aes(x = .data[[xaxis]], y = -log10(.data[[yaxis]]))) +
    geom_point(aes(color = sig), size = 3) +  
    scale_color_manual(
      name = "",
      values = c("Up" = color_up, "Down" = color_down, "Not Significant" = color_nsig),
      labels = c("Up" = uplab,  
                 "Down" = downlab, 
                 "Not Significant" = "Not Statistically Significant")) +
    labs(y = expression(bold(-Log[10](FDR-corrected_p-value))), x = expression(bold(Log[2](Fold_Change)))) +
    theme_classic(base_size = 26) +  
    theme(
      legend.direction = "vertical",
      legend.position = c(0.025, 0.95),
      legend.justification = c(0, 1),
      axis.title = element_text(size=26, face="bold"),   
      axis.text  = element_text(size=20),               
      legend.text = element_text(size=26),               
      axis.line = element_line(size = 1.0),              
      axis.ticks = element_line(size = 0.8),            
      plot.margin = margin(8, 8, 8, 8)
    ) +
    guides()
  
  if (show_threshold_lines) {
    p <- p +
      geom_hline(yintercept = -log10(thr), linetype = "dashed", color = "black", size = 0.7) +  
      geom_vline(xintercept = c(-thr2, thr2), linetype = "dashed", color = "black", size = 0.7)
  }
  
  if (label_points) {
    sig_data <- subset(inputData, sig != "Not Significant")
    p <- p +
      ggrepel::geom_text_repel(data = sig_data, aes(label = .data[[name]]),
                               size = 5,           # ↑ label font size
                               box.padding = 0.4,
                               point.padding = 0.3,
                               max.overlaps = Inf,
                               segment.color = "grey50",
                               segment.size = 0.6) 
  }
  
  if (!is.null(xlim_range)) {
    p <- p + coord_cartesian(xlim = xlim_range)
  }
  
  if (!is.null(highlight_names)) {
    highlight_data <- inputData %>% filter(.data[[name]] %in% highlight_names)
    
    p <- p +
      ggrepel::geom_text_repel(
        data = highlight_data,
        aes(label = .data[[name]]),
        size = 6,
        fontface="bold",  
        box.padding = 0.5,
        point.padding = 0.4,
        max.overlaps = Inf,
        segment.size = 0.8,         
        segment.color = "black",    
        arrow = arrow(length = unit(0.025, "npc")), 
        nudge_y = 0.3
      )
  }
  return(p)
}
```

## Function for visaulizing volcano plot 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func_2 <- function(inputData, xaxis, yaxis, uplab, downlab, name,
                               thr = 0.05, thr2 = 0, 
                               color_up = "red", color_down = "blue", color_nsig = "grey",
                               label_points = TRUE,
                               show_threshold_lines = TRUE,
                               xlim_range = NULL,
                               highlight_names=NULL) {
  
  inputData$sig <- ifelse(inputData[[yaxis]] > thr | abs(inputData[[xaxis]]) <= thr2, 
                          "Not Significant",
                          ifelse(inputData[[xaxis]] > thr2, "Up", "Down"))
  inputData$sig <- factor(inputData$sig,
                          levels = c("Up", "Down", "Not Significant"))
  
  p <- ggplot(inputData, aes(x = .data[[xaxis]], y = -log10(.data[[yaxis]]))) +
    geom_point(aes(color = sig), size = 3) +   
    scale_color_manual(
      name = "",
      values = c("Up" = color_up, "Down" = color_down, "Not Significant" = color_nsig),
      labels = c("Up" = uplab,  
                 "Down" = downlab, 
                 "Not Significant" = "Not Statistically Significant")) +
    labs(y = expression(bold(-Log[10](p-value))), x = expression(bold(Log[2](Fold_Change)))) +
    theme_classic(base_size = 26) +  
    theme(
      legend.direction = "vertical",
      legend.position = c(0.025, 0.95),
      legend.justification = c(0, 1),
      axis.title = element_text(size=26, face="bold"),   
      axis.text  = element_text(size=20),               
      legend.text = element_text(size=26),               
      axis.line = element_line(size = 1.0),             
      axis.ticks = element_line(size = 0.8),            
      plot.margin = margin(8, 8, 8, 8)
    ) +
    guides()
  
  if (show_threshold_lines) {
    p <- p +
      geom_hline(yintercept = -log10(thr), linetype = "dashed", color = "black", size = 0.7) +  
      geom_vline(xintercept = c(-thr2, thr2), linetype = "dashed", color = "black", size = 0.7)
  }
  
  if (label_points) {
    sig_data <- subset(inputData, sig != "Not Significant")
    p <- p +
      ggrepel::geom_text_repel(data = sig_data, aes(label = .data[[name]]),
                               size = 5,          
                               box.padding = 0.4,
                               point.padding = 0.3,
                               max.overlaps = Inf,
                               segment.color = "grey50",
                               segment.size = 0.6) 
  }
  
  if (!is.null(xlim_range)) {
    p <- p + coord_cartesian(xlim = xlim_range)
  }
  
  if (!is.null(highlight_names)) {
    highlight_data <- inputData %>% filter(.data[[name]] %in% highlight_names)
    
    p <- p +
      ggrepel::geom_text_repel(
        data = highlight_data,
        aes(label = .data[[name]]),
        size = 6,
        fontface="bold",  
        box.padding = 0.5,
        point.padding = 0.4,
        max.overlaps = Inf,
        segment.size = 0.8,        
        segment.color = "black",    
        arrow = arrow(length = unit(0.025, "npc")), 
        nudge_y = 0.3
      )
  }
  return(p)
}
```

## Function for visaulizing heatmap plot 

```{r}
heatmapPlot_func <- function(volcanoPlot_data, ClinicalData, MetaboData,
                             logFC_col = "logFC.TB.noTB",
                             qval_col = "qValue.TB.noTB",
                             name_col = "Metabolite",
                             id_col = "ID",
                             group_col = "TB",
                             logFC_thr = 0,
                             qval_thr = 0.05,
                             up_color = "red",
                             mid_color = "white",
                             down_color = "blue",
                             annotation_colors = c("No" = "#ca6332", "Yes" = "#3abbcc"),
                             row_fontsize = 20,
                             col_fontsize = 20) {
  
  
  Up_Metabolites <- volcanoPlot_data %>%
    filter(.data[[logFC_col]] > logFC_thr & .data[[qval_col]] < qval_thr) %>%
    arrange(.data[[qval_col]]) %>%
    pull(.data[[name_col]])
  
  Down_Metabolites <- volcanoPlot_data %>%
    filter(.data[[logFC_col]] < -logFC_thr & .data[[qval_col]] < qval_thr) %>%
    arrange(.data[[qval_col]]) %>%
    pull(.data[[name_col]])
  
  DEM <- c(Up_Metabolites, Down_Metabolites)
  
  if (length(DEM) == 0) {
    stop("No significant metabolites found with the given thresholds.")
  }
  
  sample_sorted <- ClinicalData %>%
    arrange(.data[[group_col]]) %>%
    pull(.data[[id_col]])
  
  annotation_group <- ClinicalData %>%
    arrange(.data[[group_col]]) %>%
    pull(.data[[group_col]])
  
  
  input_data <- MetaboData[DEM, sample_sorted, drop = FALSE]
  
  input_data[is.na(input_data)] <- 0
  
  input_data_scaled <- t(apply(input_data, 1, function(x) {
    s <- scale(x)
    if (all(is.na(s))) rep(0, length(x)) else s
  }))
  
  colnames(input_data_scaled) <- colnames(input_data)
  
  color_fun <- colorRamp2(c(4, 0, -4), c(up_color, mid_color, down_color))
  split_vector_rows <- factor(
    rep(c("Up", "Down"), times = c(length(Up_Metabolites), length(Down_Metabolites))),
    levels = c("Up", "Down")
  )
  
  annotation_data <- ClinicalData %>%
    arrange(.data[[group_col]]) %>%
    select(all_of(c(group_col, "Sex", "HIV")))  
  
  annotation_group <- annotation_data[[group_col]]
  annotation_sex <- annotation_data[["Sex"]]
  annotation_HIV <- annotation_data[["HIV"]]
  
  annotation_colors_list <- list(
    Group = annotation_colors,               
    Sex = c("F" = "pink", "M" = "lightblue"),
    HIV = c("Negative" = "#e24a33", "Positive" = "#9e4760")
  )
  
  
  column_annotation <- HeatmapAnnotation(
    Group = annotation_group,
    Sex = annotation_sex,
    HIV = annotation_HIV,
    col = annotation_colors_list,
    annotation_name_gp = gpar(col = "black", fontsize = 20, fontface = "bold"),
    show_legend = TRUE,
    annotation_legend_param = list(
      Group = list(
        title = "Group",
        legend_direction = "horizontal",
        legend_position = "bottom",
        at = c("Yes", "No"),
        labels = c("PTB", "ORD"),
        title_gp = gpar(fontsize = 26, fontface = "bold"),
        labels_gp = gpar(fontsize = 20)
      ),
      Sex = list(
        title = "Sex",
        legend_direction = "horizontal",
        legend_position = "bottom",
        title_gp = gpar(fontsize = 26, fontface = "bold"),
        labels_gp = gpar(fontsize = 20)
      ),
      HIV = list(
        title = "HIV status",
        legend_direction = "horizontal",
        legend_position = "bottom",
        title_gp = gpar(fontsize = 26, fontface = "bold"),
        labels_gp = gpar(fontsize = 20)
      )
    )
  )
  
  
  plot <- Heatmap(
    input_data_scaled,
    name = "Z-score",
    row_names_side = "left",
    row_names_gp = gpar(fontsize = row_fontsize, col = "black",fontface="bold"),
    row_gap = unit(1.5, "mm"),
    column_names_gp = gpar(fontsize = col_fontsize),
    cluster_columns = FALSE,
    column_gap = unit(1.5, "mm"),
    col = color_fun,
    cluster_rows = TRUE,
    show_row_dend = FALSE,
    show_column_names = FALSE,
    row_split = split_vector_rows,
    row_title = NULL,
    top_annotation = column_annotation,
    show_heatmap_legend = TRUE,
    heatmap_legend_param = list(
      title = "Z-score",
      title_gp = gpar(fontsize = 26, fontface = "bold"),
      labels_gp = gpar(fontsize = 20)
    )
  )
  
  return(plot)
}

```

## Function for visaulizing PCA plot

```{r}
PCAplot <- function(pca.result, sampleData){
  inputData <- as.data.frame(pca.result$x)
  inputData$TB <- factor(sampleData$TB)
  inputData$HIV <- factor(sampleData$HIV)
  
  plot <- ggplot(inputData, aes(x=PC1, y=PC2)) +
    geom_point(size=5, aes(shape=HIV, color=TB), stroke=1) +
    stat_ellipse(aes(color=TB), type="t", linetype=2) +
    scale_color_manual(values=c("Yes"="#3abbcc","No"= "#ca6332"),labels = c("Yes"="PTB", "No"="ORD"), name="Group") +
    labs(
      shape = "HIV status",
      x = paste0("PC1 (", round(pca.result$importance[2,1] * 100, 1), "%)"),
      y = paste0("PC2 (", round(pca.result$importance[2,2] * 100, 1), "%)")
    ) +
    theme_classic(base_size = 26) +
    theme(
      axis.title = element_text(size=26, face="bold"),
      axis.text  = element_text(size=20),
      legend.position = "top",
      legend.title = element_text(size=26, face="bold"),
      legend.text  = element_text(size=26),
      axis.line = element_line(size = 0.8),   
      axis.ticks = element_line(size = 0.6)   
    )
  
  return(plot)
}
```

## Function for visualizing ROC plot for test using combined, cytokine and metabolite

```{r}
threerocPlot_func <- function(roc1, roc2, roc3,
                              names = c("Combined", "Cytokine", "Metabolite"),
                              colors = c("#7aa456", "#c65999", "#777acd")) {
  
  if (length(names) != 3) stop("Please provide exactly 3 names")
  if (length(colors) != 3) stop("Please provide exactly 3 colors")
  
  ci1 <- ci.auc(roc1)
  ci2 <- ci.auc(roc2)
  ci3 <- ci.auc(roc3)
  
  auc1 <- sprintf("%s AUC = %.2f (%.2f–%.2f)", names[1], auc(roc1), ci1[1], ci1[3])
  auc2 <- sprintf("%s AUC = %.2f (%.2f–%.2f)", names[2], auc(roc2), ci2[1], ci2[3])
  auc3 <- sprintf("%s AUC = %.2f (%.2f–%.2f)", names[3], auc(roc3), ci3[1], ci3[3])
  
  sens1 <- ci.coords(roc1, x = 0.98, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens2 <- ci.coords(roc2, x = 0.98, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens3 <- ci.coords(roc3, x = 0.98, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  
  sens1_triage <- ci.coords(roc1, x = 0.70, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens2_triage <- ci.coords(roc2, x = 0.70, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens3_triage <- ci.coords(roc3, x = 0.70, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  
  roc_list <- setNames(
    list(roc1, roc2, roc3),
    c(auc1, auc2, auc3)
  )
  
  plot <- ggroc(roc_list, size = 2) +   
    theme_classic(base_size = 28) +      
    theme(
      legend.position = c(0.55, 0.25),
      legend.title = element_blank(),
      legend.text = element_text(size = 26),      
      axis.title.x = element_text(size = 26, face="bold"), 
      axis.title.y = element_text(size = 26, face="bold"),
      axis.text = element_text(size = 20),        
      plot.title = element_text(hjust = 0.5, size=26, face="bold")
    ) +
    scale_color_manual(values = colors)
  
  return(list(
    plot = plot,
    sensitivity = setNames(list(sens1, sens2, sens3), names),
    sensitivity_triage = setNames(list(sens1_triage, sens2_triage, sens3_triage), names)
  ))
}
```

## Function for manhattan plot  
```{r}
manhattan_plot <- function(data, 
                           x_axis = "m/z", 
                           pval_col = "p-value", 
                           sig_col = "sig",
                           alpha_threshold = 0.0001,
                           sig_colors = c("Significant" = "red", "Not Significant" = "grey70"),
                           point_size = 2,
                           axis_title_size = 26,
                           axis_text_size = 26) {
  
  
  data$logP <- -log10(data[[pval_col]])
  
  data[[sig_col]] <- ifelse(data[[pval_col]] < alpha_threshold , "Significant", "Not Significant")
  
  plot <- ggplot(data, aes(x = .data[[x_axis]], y = logP, color = .data[[sig_col]])) +
    geom_point(size = point_size, alpha = 0.8) +
    scale_color_manual(values = sig_colors) +
    geom_hline(yintercept = -log10(alpha_threshold), linetype = "dashed", 
               color = "blue", size = 0.8) +
    labs(x = x_axis, y = expression(bold(-log[10](p-value))), color = "") +
    theme_classic(base_size = axis_text_size) +
    theme(
      axis.title = element_text(size = axis_title_size, face = "bold"),
      axis.text = element_text(size = axis_text_size),
      legend.position = "top",
      legend.text = element_text(size = 26),
      plot.margin = margin(5, 5, 5, 5)
    )
  
  return(plot)
}
```

## Function for visualizing ROC plot for test using combined, HIV pos and HIV neg

```{r}
threerocPlot_func_HIV <- function(roc_test, roc_test_HIVpos, roc_test_HIVneg) {
  
  ci_test <- ci.auc(roc_test)
  ci_HIVpos <- ci.auc(roc_test_HIVpos)
  ci_HIVneg <- ci.auc(roc_test_HIVneg)
  
  auc_test <- sprintf("Combined AUC = %.2f (%.2f–%.2f)", auc(roc_test), ci_test[1], ci_test[3])
  auc_HIVpos <- sprintf("HIV+ AUC = %.2f (%.2f–%.2f)", auc(roc_test_HIVpos), ci_HIVpos[1], ci_HIVpos[3])
  auc_HIVneg <- sprintf("HIV− AUC = %.2f (%.2f–%.2f)", auc(roc_test_HIVneg), ci_HIVneg[1], ci_HIVneg[3])
  
  sens_test <- ci.coords(roc_test, x = 0.98, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens_HIVpos <- ci.coords(roc_test_HIVpos, x = 0.98, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens_HIVneg <- ci.coords(roc_test_HIVneg, x = 0.98, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  
  sens_test_triage <- ci.coords(roc_test, x = 0.80, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens_HIVpos_triage <- ci.coords(roc_test_HIVpos, x = 0.80, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  sens_HIVneg_triage <- ci.coords(roc_test_HIVneg, x = 0.80, input = "specificity", ret = "sensitivity", conf.level = 0.95)
  
  roc_list <- setNames(
    list(roc_test, roc_test_HIVpos, roc_test_HIVneg),
    c(auc_test, auc_HIVpos, auc_HIVneg)
  )
  
  plot <- ggroc(roc_list, size = 1) +
    theme_classic(base_size = 26) +
    theme(
      legend.position = c(0.6, 0.4),
      legend.title = element_blank(),
      plot.title = element_text(hjust = 0.5),
      axis.title.x = element_text(size = 20),
      axis.title.y = element_text(size = 20) 
    ) +
    scale_color_manual(values = c("#7aa456", "#D55E00", "violetred4"))
  
  sensitivity_results <- list(
    Combined = sens_test,
    HIVpos = sens_HIVpos,
    HIVneg = sens_HIVneg
  )
  
  sensitivity_triage_results <- list(
    Combined = sens_test_triage,
    HIVpos = sens_HIVpos_triage,
    HIVneg = sens_HIVneg_triage
  )
  
  return(list(
    plot = plot,
    sensitivity = sensitivity_results,
    sensitivity_triage = sensitivity_triage_results
  ))
}
```

## Function for visualizing ROC plot for both train and test

```{r}
bothrocPlot_func <- function(roc_train,roc_test){
  
  auc_train <- round(auc(roc_train), 3)
  auc_test <- round(auc(roc_test), 3)
  
  roc_list <- setNames(
    list(roc_train, roc_test),
    c(paste0("Training Set AUC = ", auc_train),
      paste0("Test Set AUC = ", auc_test))
  )
  
  plot <- ggroc(roc_list, size = 1) +
    geom_hline(yintercept = 0.9, linetype = "dashed", color = "gray40") +
    geom_vline(xintercept = 0.7, linetype = "dashed", color = "gray40") +
    geom_rect(aes(xmin = 0.7, xmax = 1, ymin = 0.9, ymax = 1),
              fill = "gray85", alpha = 0.4, inherit.aes = FALSE) +
    theme_classic(base_size = 18) +
    theme(
      legend.position = c(0.6, 0.4),
      legend.title = element_blank(),
      plot.title = element_text(hjust = 0.5)
    ) +
    scale_color_manual(values = c("#D55E00", "violetred4"))
  return(plot)
}
```


## Function for visualizing ROC plot for both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func <- function(roc_train,roc_test,roc_test_HIVpos,roc_test_HIVneg){
  
  auc_train <- round(auc(roc_train), 2)
  auc_test <- round(auc(roc_test), 2)
  auc_test_HIVpos <- round(auc(roc_test_HIVpos), 2)
  auc_test_HIVneg <- round(auc(roc_test_HIVneg), 2)
  
  roc_list <- setNames(
    list(roc_train, roc_test,roc_test_HIVpos,roc_test_HIVneg),
    c(paste0("Training Set AUC = ", auc_train),
      paste0("Test Set AUC = ", auc_test),
      paste0("Test HIV+ Set AUC = ", auc_test_HIVpos),
      paste0("Test HIV- Set AUC = ", auc_test_HIVneg))
  )
  
  plot <- ggroc(roc_list, size = 1) +
    geom_hline(yintercept = 0.9, linetype = "dashed", color = "gray40") +
    geom_vline(xintercept = 0.7, linetype = "dashed", color = "gray40") +
    geom_rect(aes(xmin = 0.7, xmax = 1, ymin = 0.9, ymax = 1),
              fill = "gray85", alpha = 0.4, inherit.aes = FALSE) +
    theme_classic(base_size = 18) +
    theme(
      legend.position = c(0.6, 0.4),
      legend.title = element_blank(),
      plot.title = element_text(hjust = 0.5)
    ) +
    scale_color_manual(values = c("#D55E00", "violetred4","darkgreen","darkblue"))
  return(plot)
}
```

## Function for visualizing ROC plot for both train and test and both no TB SA and no TB TBGWAS

```{r}
fourrocPlot_func_2 <- function(roc_train,roc_test,roc_test_noTBSA,roc_test_noTBGWAS){
  
  auc_train <- round(auc(roc_train), 2)
  auc_test <- round(auc(roc_test), 2)
  auc_test_noTBSA <- round(auc(roc_test_noTBSA), 2)
  auc_test_noTBGWAS <- round(auc(roc_test_noTBGWAS), 2)
  
  roc_list <- setNames(
    list(roc_train, roc_test,roc_test_noTBSA,roc_test_noTBGWAS),
    c(paste0("Training Set AUC = ", auc_train),
      paste0("Test Set AUC = ", auc_test),
      paste0("Test noTB SA Set AUC = ", auc_test_noTBSA),
      paste0("Test noTB TBGWAS Set AUC = ", auc_test_noTBGWAS))
  )
  
  plot <- ggroc(roc_list, size = 1) +
    geom_hline(yintercept = 0.9, linetype = "dashed", color = "gray40") +
    geom_vline(xintercept = 0.7, linetype = "dashed", color = "gray40") +
    geom_rect(aes(xmin = 0.7, xmax = 1, ymin = 0.9, ymax = 1),
              fill = "gray85", alpha = 0.4, inherit.aes = FALSE) +
    theme_classic(base_size = 18) +
    theme(
      legend.position = c(0.6, 0.4),
      legend.title = element_blank(),
      plot.title = element_text(hjust = 0.5)
    ) +
    scale_color_manual(values = c("#D55E00", "violetred4","darkgreen","darkblue"))
  return(plot)
}
```


## Function for boxplot visualization

```{r}
boxplot_func <- function(data, xaxis, yaxis, group){
  y_min <- min(data$value, na.rm = TRUE) - 0.2
  y_max <- max(data$value, na.rm = TRUE) + 0.2
  
  plot <- ggplot(data, aes(x = .data[[xaxis]], y = .data[[yaxis]], fill = .data[[group]])) +
    geom_boxplot(outlier.shape = NA, alpha = 0.6, width = 0.6, 
                 position = position_dodge(width = 0.75)) +
    geom_jitter(aes(color = .data[[group]]), 
                position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.75),
                size = 2.5, alpha = 1) +
    scale_color_manual(values = c("Yes"="#3abbcc","No"= "#ca6332"),
                       labels = c("Yes"="PTB","No"= "ORD")) +
    scale_fill_manual(values = c("Yes"="#3abbcc","No"= "#ca6332")) +
    labs(y = expression(bold(log[2](Plasma_Concentration))), x = NULL, color = NULL) +
    guides(fill = "none") +
    theme_classic(base_size = 20) +
    theme(
      axis.title = element_text(size = 30, face = "bold"),
      axis.text  = element_text(size = 30, face="bold",angle = 45, hjust = 1),
      legend.position = "right",
      legend.title = element_text(size = 30, face = "bold"),
      legend.text  = element_text(size = 30),
      axis.line = element_line(size = 0.9),
      axis.ticks = element_line(size = 0.7)
    ) +
    coord_cartesian(ylim = c(y_min, y_max)) +
    
    stat_compare_means(
      aes(group = .data[[group]]),
      method = "wilcox.test",      
      label = "p.signif",        
      hide.ns = TRUE,             
      label.y = y_max + 0.1,
      size = 6,                  
      fontface = "bold"  
    )
  
  return(plot)
}
```
# Which data? {.tabset}
## Data at baseline (only adults) - PCA Metabolites

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
MetaboData_Baseline_Adult <- MetaboData_Baseline_Adult[ClinicalData_Baseline_Adult$ID,]
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
ClinicalData_Baseline_Adult$Group <- cbind(paste0(ClinicalData_Baseline_Adult$TB,",",ClinicalData_Baseline_Adult$HIV))
dim(MetaboData_Baseline_Adult)
```


### Creating a covariates matrix

```{r}
pca.metabolite <- prcomp(MetaboData_Baseline_Adult, scale. = TRUE,center = TRUE)
plot(pca.metabolite)
pca.result <- summary(pca.metabolite)
Fig1_a_pca <- PCAplot(pca.result, ClinicalData_Baseline_Adult)
```


## Data at baseline (only adults) - PCA Cyotkines

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult[match(row.names(CytokineData_Baseline_Adult_imputed),ClinicalData_Baseline_Adult$ID),]
dim(CytokineData_Baseline_Adult)
```


### Creating a covariates matrix

```{r}
pca.cytokine <- prcomp(CytokineData_Baseline_Adult_imputed, scale. = TRUE,center = TRUE)
plot(pca.cytokine)
pca.result <- summary(pca.cytokine)
Fig1_b_pca <- PCAplot(pca.result, ClinicalData_Baseline_Adult)
```


## Data at baseline (only adults) - DE Metabolites

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
MetaboData_Baseline_Adult <- MetaboData_Baseline_Adult[ClinicalData_Baseline_Adult$ID,]
dim(MetaboData_Baseline_Adult)
```


### Creating a covariates matrix

```{r}
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()%>%t()
covareiateData <- ClinicalData_Baseline_Adult[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult$TB)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("noTB","TB","Age","HIV","Sex")
cont.matrix <- makeContrasts(TB-noTB,levels=design)
fit <- lmFit(MetaboData_Baseline_Adult, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "fdr"))
volcanoPlot_data_MET <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_MET) <- c("logFC.TB.noTB","pValue.TB.noTB","qValue.TB.noTB")

volcanoPlot_data_MET$Metabolite <- row.names(volcanoPlot_data_MET)
save(volcanoPlot_data_MET,file="../Results/DEA/DEM_Baseline_Adult.RData")
volcanoPlot_data_MET%>%
  filter(logFC.TB.noTB>0&pValue.TB.noTB<0.05)%>%
  nrow()
volcanoPlot_data_MET%>%
  filter(logFC.TB.noTB<0&pValue.TB.noTB<0.05)%>%
  nrow()
volcanoPlot_data_MET%>%
  filter(logFC.TB.noTB>0&qValue.TB.noTB<0.05)%>%
  nrow()
volcanoPlot_data_MET%>%
  filter(logFC.TB.noTB<0&qValue.TB.noTB<0.05)%>%
  nrow()
```

### Volcano plot for TB vs. no TB 

#### Using p-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func_2(volcanoPlot_data_MET, 
                   xaxis = "logFC.TB.noTB", 
                   yaxis = "pValue.TB.noTB", 
                   uplab = "Up in PTB", 
                   downlab = "Down in PTB", 
                   name = "Metabolite",
                   thr = 0.05, thr2 = 0, 
                   label_points = FALSE,
                   xlim_range = c(-3, 3),
                   highlight_names = c("METHIONINE", "OXOPROLINE"))
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
Fig1_a_volcano <- volcanoPlot_func(volcanoPlot_data_MET, 
                                   xaxis = "logFC.TB.noTB", 
                                   yaxis = "qValue.TB.noTB", 
                                   uplab = "Up in PTB", 
                                   downlab = "Down in PTB", 
                                   name = "Metabolite",
                                   thr = 0.05, thr2 = 0, 
                                   label_points = FALSE,
                                   xlim_range = c(-3, 3),
                                   highlight_names = c("METHIONINE", "OXOPROLINE"))
```

### Visualizing heatmap plot 

```{r}
myHeatmap <- heatmapPlot_func(volcanoPlot_data_MET, ClinicalData_Baseline_Adult, MetaboData_Baseline_Adult,name_col = "Metabolite")
draw(myHeatmap)
Fig1_a_heatmap <- grid.grabExpr(draw(myHeatmap, newpage = FALSE))
```


## Data at baseline (only adults) - DE Cytokines

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult[match(row.names(CytokineData_Baseline_Adult),ClinicalData_Baseline_Adult$ID),]
dim(CytokineData_Baseline_Adult)
```


### Creating a covariates matrix

```{r}
CytokineData_Baseline_Adult_imputed <- CytokineData_Baseline_Adult_imputed%>%t()
covareiateData <- ClinicalData_Baseline_Adult[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult$TB)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("noTB","TB","Age","HIV","Sex")
cont.matrix <- makeContrasts(TB-noTB,levels=design)
fit <- lmFit(CytokineData_Baseline_Adult_imputed, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "BH"))
volcanoPlot_data_CYT <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_CYT) <- c("logFC.TB.noTB","pValue.TB.noTB","qValue.TB.noTB")

volcanoPlot_data_CYT$Cytokine <- row.names(volcanoPlot_data_CYT)
save(volcanoPlot_data_CYT,file="../Results/DEA/DEC_Baseline_Adult.RData")
volcanoPlot_data_CYT%>%
  filter(logFC.TB.noTB>0&pValue.TB.noTB<0.05)%>%
  nrow()
volcanoPlot_data_CYT%>%
  filter(logFC.TB.noTB<0&pValue.TB.noTB<0.05)%>%
  nrow()
volcanoPlot_data_CYT%>%
  filter(logFC.TB.noTB>0&qValue.TB.noTB<0.05)%>%
  nrow()
volcanoPlot_data_CYT%>%
  filter(logFC.TB.noTB<0&qValue.TB.noTB<0.05)%>%
  nrow()
```

### Volcano plot for TB vs. no TB 

#### Using p-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func_2(volcanoPlot_data_CYT, 
                   xaxis = "logFC.TB.noTB", 
                   yaxis = "pValue.TB.noTB", 
                   uplab = "Up in PTB", 
                   downlab = "Down in PTB", 
                   name = "Cytokine",
                   thr = 0.05, thr2 = 0, 
                   label_points = FALSE,
                   xlim_range = c(-4, 4),
                   highlight_names = c("IFN.gamma" ,"IL.1beta" , "IL.22"   ,  "IL.10"   ,  "IFN.beta"))
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
Fig1_b_volcano <- volcanoPlot_func(volcanoPlot_data_CYT, 
                                   xaxis = "logFC.TB.noTB", 
                                   yaxis = "qValue.TB.noTB", 
                                   uplab = "Up in PTB", 
                                   downlab = "Down in PTB", 
                                   name = "Cytokine",
                                   thr = 0.05, thr2 = 0, 
                                   label_points = FALSE,
                                   xlim_range = c(-4, 4),
                                   highlight_names = c("IFN.gamma" ,"IL.1beta" , "IL.22"   ,  "IL.10"   ,  "IFN.beta"))
```

### Visualizing heatmap plot 

```{r}
myHeatmap <- heatmapPlot_func(volcanoPlot_data_CYT, ClinicalData_Baseline_Adult, CytokineData_Baseline_Adult_imputed,name_col="Cytokine")
draw(myHeatmap)
Fig1_b_heatmap <- grid.grabExpr(draw(myHeatmap, newpage = FALSE))
```

## Make Supplementary Figures 1 and 2

```{r}
SuppFig1 <- plot_grid((Fig1_a_pca/Fig1_a_volcano),Fig1_a_heatmap, ncol = 2, rel_widths = c(1, 1.25))
SuppFig2 <- plot_grid((Fig1_b_pca/Fig1_b_volcano),Fig1_b_heatmap, ncol = 2, rel_widths = c(1, 1.25))
pdf(file="../Results/Figures/Manuscript/SuppFig1.pdf",height = 22,width = 32)
plot_grid(SuppFig1,nrow=1,ncol=1,rel_heights = c(1,1))
dev.off()
pdf(file="../Results/Figures/Manuscript/SuppFig2.pdf",height = 22,width = 32)
plot_grid(SuppFig2,nrow=1,ncol=1,rel_heights = c(1,1))
dev.off()
```



## Data at baseline (only adults) - Multiomics biomarker
### Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()
```

### Make train and test data  


```{r message=FALSE, warning=FALSE}
ClinicalData_Baseline_Adult$Group <- cbind(paste0(ClinicalData_Baseline_Adult$TB,",",ClinicalData_Baseline_Adult$HIV))

set.seed(123) 
test_percentage <- 0.3 

test_indices <- createDataPartition(ClinicalData_Baseline_Adult$Group, p = test_percentage, list = FALSE)

sampleData_train <- ClinicalData_Baseline_Adult[-test_indices, ]
sampleData_test <- ClinicalData_Baseline_Adult[test_indices, ]


removed_samles_train <- which(!sampleData_train$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_train <- sampleData_train[-removed_samles_train,]
removed_samles_test <- which(!sampleData_test$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_test <- sampleData_test[-removed_samles_test,]

table(sampleData_train$TB,sampleData_train$HIV)
table(sampleData_test$TB,sampleData_test$HIV)

metaboData.train <- MetaboData_Baseline_Adult[sampleData_train$ID,]
cytokineData.train <- CytokineData_Baseline_Adult_imputed[sampleData_train$ID,]
sampleData.train <- sampleData_train
metaboData.test <- MetaboData_Baseline_Adult[sampleData_test$ID,]
cytokineData.test <- CytokineData_Baseline_Adult_imputed[sampleData_test$ID,]
sampleData.test <- sampleData_test


metaboData.train <- as.data.frame(metaboData.train)
cytokineData.train <- as.data.frame(cytokineData.train)

sampleData_test_HIVpos <- sampleData_test%>%
  filter(HIV=="Positive")
sampleData_test_HIVneg <- sampleData_test%>%
  filter(HIV=="Negative")
sampleData_test_noTBSA <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "SA"))
sampleData_test_noTBGWAS <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "TBGWAS"))

metaboData.test.HIVpos <- metaboData.test[sampleData_test_HIVpos$ID,]
cytokineData.test.HIVpos <- cytokineData.test[sampleData_test_HIVpos$ID,]
sampleData.test.HIVpos <- sampleData_test_HIVpos

metaboData.test.HIVneg <- metaboData.test[sampleData_test_HIVneg$ID,]
cytokineData.test.HIVneg <- cytokineData.test[sampleData_test_HIVneg$ID,]
sampleData.test.HIVneg <- sampleData_test_HIVneg

metaboData.test.noTBSA <- metaboData.test[sampleData_test_noTBSA$ID,]
cytokineData.test.noTBSA <- cytokineData.test[sampleData_test_noTBSA$ID,]
sampleData.test.noTBSA <- sampleData_test_noTBSA

metaboData.test.noTBGWAS <- metaboData.test[sampleData_test_noTBGWAS$ID,]
cytokineData.test.noTBGWAS <- cytokineData.test[sampleData_test_noTBGWAS$ID,]
sampleData.test.noTBGWAS <- sampleData_test_noTBGWAS
```

### Make a combined data including metabolites and cytokine for feature selection

```{r}
combinedData.train <- cbind(metaboData.train,cytokineData.train)
combinedData.test <- cbind(metaboData.test,cytokineData.test)
combinedData.test.HIVpos <- cbind(metaboData.test.HIVpos,cytokineData.test.HIVpos)
combinedData.test.HIVneg <- cbind(metaboData.test.HIVneg,cytokineData.test.HIVneg)
combinedData.test.noTBSA <- cbind(metaboData.test.noTBSA,cytokineData.test.noTBSA)
combinedData.test.noTBGWAS <- cbind(metaboData.test.noTBGWAS,cytokineData.test.noTBGWAS)
combinedData.train$Group= as.factor(sampleData.train$TB)
combinedData.train$Group <- factor(combinedData.train$Group, levels =c("Yes","No"))
```

### Run Boruta

```{r eval=FALSE}
set.seed(111)
boruta_data <- Boruta(Group~., data = combinedData.train, doTrace = 2, maxRuns=400)
boruta_data <- TentativeRoughFix(boruta_data)
boruta_stats <- attStats(boruta_data)
boruta_stats$cytokine <- rownames(boruta_stats)
boruta_stats <- boruta_stats %>% filter(decision =="Confirmed")
boruta_stats <- boruta_stats[order(boruta_stats$meanImp,decreasing = TRUE),c("cytokine","meanImp")]
plot(boruta_data, cex.axis = 1.1, xlab = "", las = 2,ylab="")
title("Cytokine",line=+1,adj=0,cex.main=2)
combined_boruta <- boruta_stats$cytokine
combined_boruta
combined_boruta_top3 <- combined_boruta[1:3]
combined_boruta_top5 <- combined_boruta[1:5]
saveRDS(combined_boruta,file="../Results/Boruta/selectedCombined_baseline_adult.rds")
saveRDS(combined_boruta_top3, file="../Results/Boruta/selectedCombined_top3_baseline_adult.rds")
saveRDS(combined_boruta_top5, file="../Results/Boruta/selectedCombined_top5_baseline_adult.rds")
combined_boruta <- readRDS("../Results/Boruta/selectedCombined_baseline_adult.rds")
```

### Run randomforestSRC

```{r eval=FALSE}
set.seed(111)
rf_model <- rfsrc(Group ~ ., data = combinedData.train, ntree = 500, importance = TRUE)
vimp <- rf_model$importance
vimp[order(vimp[,1],decreasing = TRUE),]
vimp.rfsrc(rf_model)
min_depth <- var.select(rf_model)
selected_combined_SRC <- min_depth$md.obj$topvars
selected_combined_SRC_top2 <- head(min_depth$topvars, 2)
selected_combined_SRC_top5 <- head(min_depth$topvars, 5)
selected_combined_SRC_top7 <- head(min_depth$topvars, 7)
saveRDS(selected_combined_SRC,file="../Results/randomforestSRC/selectedCombined_baseline_adult.rds")
saveRDS(selected_combined_SRC_top2, file="../Results/randomforestSRC/selectedCombined_top2_baseline_adult.rds")
saveRDS(selected_combined_SRC_top5, file="../Results/randomforestSRC/selectedCombined_top5_baseline_adult.rds")
saveRDS(selected_combined_SRC_top7, file="../Results/randomforestSRC/selectedCombined_top7_baseline_adult.rds")
combined_boruta <- readRDS("../Results/randomforestSRC/selectedCombined_baseline_adult.rds")
```

### Selected cytokines from randomforestSRC

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("IFN.gamma" ,"IL.1beta" , "IL.22"   ,  "IL.10"   ,  "IFN.beta")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_cytokine_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_cytokine_test_noTBSA <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_cytokine_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_noTBGWAS <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train,roc_cytokine_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train,roc_cytokine_test,roc_cytokine_test_HIVpos,roc_cytokine_test_HIVneg)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_cytokine_train,roc_cytokine_test,roc_cytokine_test_noTBSA,roc_cytokine_test_noTBGWAS)
```


### Selected metabolites from randomforestSRC

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_metabols <- c("METHIONINE", "OXOPROLINE")
selected_metabols

data <-as.data.frame(metaboData.train[,selected_metabols,drop=FALSE])
data$Class = as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(metaboData.test[,selected_metabols,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(metaboData.test.HIVpos[,selected_metabols,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(metaboData.test.HIVneg[,selected_metabols,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(metaboData.test.noTBSA[,selected_metabols,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(metaboData.test.noTBGWAS[,selected_metabols,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```


#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_metabol_train <- predict(svmTune, data)
prob_metabol_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_metabol_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_metabol_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_metabol_test <- predict(svmTune, data_test)
prob_metabol_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_metabol_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_metabol_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_metabol_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVneg <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM model

```{r}
pred_metabol_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_metabol_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class, 
            predictor = prob_metabol_test_noTBSA[,1], 
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_noTBSA <- roc0
```










##### Entire test data (only no TB TBGWAS)
###### Test SVM svmTune 

```{r}
pred_metabol_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_metabol_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class, 
            predictor = prob_metabol_test_noTBGWAS[,1], 
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_noTBGWAS <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_metabol_train,roc_metabol_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_metabol_train,roc_metabol_test,roc_metabol_test_HIVpos,roc_metabol_test_HIVneg)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_metabol_train,roc_metabol_test,roc_metabol_test_noTBSA,roc_metabol_test_noTBGWAS)
```



### Combined model

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <- c("IFN.gamma" ,"IL.1beta" , "IL.22"   ,  "IL.10"   ,  "IFN.beta")
selected_metabols <- selected_metabols <- c("METHIONINE", "OXOPROLINE")
selected_combined <- c("IFN.gamma" ,"IL.1beta" , "IL.22"   ,  "IL.10"   ,  "IFN.beta","METHIONINE", "OXOPROLINE")

data <- as.data.frame(cbind(cytokineData.train[,selected_cytokines,drop=FALSE],metaboData.train[,selected_metabols,drop=FALSE]))
data$Class = as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cbind(cytokineData.test[,selected_cytokines,drop=FALSE],metaboData.test[,selected_metabols,drop=FALSE]))
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cbind(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE],metaboData.test.HIVpos[,selected_metabols,drop=FALSE]))
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cbind(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE],metaboData.test.HIVneg[,selected_metabols,drop=FALSE]))
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cbind(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE],metaboData.test.noTBSA[,selected_metabols,drop=FALSE]))
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cbind(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE],metaboData.test.noTBGWAS[,selected_metabols,drop=FALSE]))
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_combined_train <- predict(svmTune, data)
prob_combined_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_combined_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_combined_test <- predict(svmTune, data_test)
prob_combined_test <- predict(svmTune, data_test, type="prob")

ptb_prob <- prob_combined_test[, "Yes"]
cor.test(ptb_prob, sampleData.test$AGE, method="spearman")

df <- sampleData.test
df$Score <- ptb_prob
lm(Score ~ TB + AGE + Sex + HIV, data = df)
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_combined_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_combined_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_combined_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_combined_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_combined_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_combined_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_combined_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_combined_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_HIVneg <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_combined_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_combined_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_combined_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_combined_test_noTBSA <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_combined_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_combined_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_combined_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_noTBGWAS <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_combined_train,roc_combined_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_combined_train,roc_combined_test,roc_combined_test_HIVpos,roc_combined_test_HIVneg)
result <- threerocPlot_func_HIV(roc_combined_test,roc_combined_test_HIVpos,roc_combined_test_HIVneg)
result$plot
SuppFig3 <- result$plot
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_combined_train,roc_combined_test,roc_combined_test_noTBSA,roc_combined_test_noTBGWAS)
```



### Genarate figures
```{r}
result <- threerocPlot_func(roc_combined_test, roc_cytokine_test, roc_metabol_test)
result$plot  
SuppFig3_total <- result$plot 
result <- threerocPlot_func(roc_combined_test, roc_combined_test_HIVpos, roc_combined_test_HIVneg, names=c("Combined","HIV-positive","HIV-negative"),colors=c("#50b957", "red","#d89f32"))
result$plot  
SuppFig3_combined_HIV <- result$plot 
result <- threerocPlot_func(roc_combined_test, roc_combined_test_noTBGWAS, roc_combined_test_noTBSA,names=c("Combined","Symptomatic contacts","Hospitalized ORD"),colors=c("#50b957", "orange","#729cda"))
result$plot  
SuppFig3_combined_noTB <- result$plot 
result <- threerocPlot_func(roc_cytokine_test, roc_cytokine_test_HIVpos, roc_cytokine_test_HIVneg,names=c("Cytokine","HIV-positive","HIV-negative"),colors=c("#d84882",  "red","#d89f32"))
result$plot  
SuppFig3_cytokine_HIV <- result$plot 
result <- threerocPlot_func(roc_cytokine_test, roc_cytokine_test_noTBGWAS, roc_cytokine_test_noTBSA,names=c("Cytokine","Symptomatic contacts","Hospitalized ORD"),colors=c("#d84882","orange" ,"#729cda"))
result$plot  
SuppFig3_cytokine_noTB <- result$plot 
result <- threerocPlot_func(roc_metabol_test, roc_metabol_test_HIVpos, roc_metabol_test_HIVneg,
                            names=c("Metabolite","HIV-positive","HIV-negative"),colors=c("#7b63d4", "red","#d89f32"))
result$plot  
SuppFig3_metabol_HIV <- result$plot 
result <- threerocPlot_func(roc_metabol_test, roc_metabol_test_noTBGWAS, roc_metabol_test_noTBSA,names=c("Metabolite","Symptomatic contacts","Hospitalized ORD"),colors=c("#7b63d4", "orange","#729cda"))
result$plot  
SuppFig3_metabol_noTB <- result$plot 

save(SuppFig3_metabol_noTB,SuppFig3_metabol_HIV,SuppFig3_cytokine_noTB,SuppFig3_cytokine_HIV,SuppFig3_combined_noTB,SuppFig3_combined_HIV,SuppFig3_total,file="../Results/SuppFig3.RData")
```

## Make Supplementary Figure 3
```{r}
l <- load("../Results/SuppFig3.RData")

SuppFig3_total$layers <-SuppFig3_total$layers[!sapply(SuppFig3_total$layers, function(x) inherits(x$geom, "GeomVline"))]
tpp_boxes <- data.frame(
  xmin = c(0.7, 0.98),   # Specificity lower limits
  xmax = c(1.0, 1.0),   # Specificity upper limits
  ymin = c(0.9, 0.65),  # Sensitivity lower limits
  ymax = c(1.0, 1),   # Sensitivity upper limits
  label = c("Triage Test", "POC Diagnostic")
)
SuppFig3_total <- SuppFig3_total+ geom_rect(data = tpp_boxes,
                                            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = label),
                                            alpha = 0.3, inherit.aes = FALSE) +
  scale_fill_manual(values = c("Triage Test" = "green", "POC Diagnostic" = "red"))+
  guides(color = guide_legend(title = "TPP Boxes"))+
  labs(title = "ROC Curves with Target Product Profile (TPP) Thresholds")+
  theme(title = element_text(size=12))


pdf(file="../Results/Figures/Manuscript/SuppFig3.pdf",width=10,height = 10)
SuppFig3_total
dev.off()
```

## Data at baseline (only adults) - 5-marker Multi-omics biomarker
### Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()
```

### Make train and test data  


```{r message=FALSE, warning=FALSE}
ClinicalData_Baseline_Adult$Group <- cbind(paste0(ClinicalData_Baseline_Adult$TB,",",ClinicalData_Baseline_Adult$HIV))

set.seed(123) 
test_percentage <- 0.3 

test_indices <- createDataPartition(ClinicalData_Baseline_Adult$Group, p = test_percentage, list = FALSE)

sampleData_train <- ClinicalData_Baseline_Adult[-test_indices, ]
sampleData_test <- ClinicalData_Baseline_Adult[test_indices, ]


removed_samles_train <- which(!sampleData_train$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_train <- sampleData_train[-removed_samles_train,]
removed_samles_test <- which(!sampleData_test$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_test <- sampleData_test[-removed_samles_test,]

table(sampleData_train$TB,sampleData_train$HIV)
table(sampleData_test$TB,sampleData_test$HIV)

metaboData.train <- MetaboData_Baseline_Adult[sampleData_train$ID,]
cytokineData.train <- CytokineData_Baseline_Adult_imputed[sampleData_train$ID,]
sampleData.train <- sampleData_train
metaboData.test <- MetaboData_Baseline_Adult[sampleData_test$ID,]
cytokineData.test <- CytokineData_Baseline_Adult_imputed[sampleData_test$ID,]
sampleData.test <- sampleData_test


metaboData.train <- as.data.frame(metaboData.train)
cytokineData.train <- as.data.frame(cytokineData.train)

sampleData_test_HIVpos <- sampleData_test%>%
  filter(HIV=="Positive")
sampleData_test_HIVneg <- sampleData_test%>%
  filter(HIV=="Negative")
sampleData_test_noTBSA <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "SA"))
sampleData_test_noTBGWAS <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "TBGWAS"))

metaboData.test.HIVpos <- metaboData.test[sampleData_test_HIVpos$ID,]
cytokineData.test.HIVpos <- cytokineData.test[sampleData_test_HIVpos$ID,]
sampleData.test.HIVpos <- sampleData_test_HIVpos

metaboData.test.HIVneg <- metaboData.test[sampleData_test_HIVneg$ID,]
cytokineData.test.HIVneg <- cytokineData.test[sampleData_test_HIVneg$ID,]
sampleData.test.HIVneg <- sampleData_test_HIVneg

metaboData.test.noTBSA <- metaboData.test[sampleData_test_noTBSA$ID,]
cytokineData.test.noTBSA <- cytokineData.test[sampleData_test_noTBSA$ID,]
sampleData.test.noTBSA <- sampleData_test_noTBSA

metaboData.test.noTBGWAS <- metaboData.test[sampleData_test_noTBGWAS$ID,]
cytokineData.test.noTBGWAS <- cytokineData.test[sampleData_test_noTBGWAS$ID,]
sampleData.test.noTBGWAS <- sampleData_test_noTBGWAS
```

### Selected cytokines in the 5-marker signature

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("IFN.gamma" , "IL.22"   ,  "IL.10")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_cytokine_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_cytokine_test_noTBSA <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_cytokine_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_noTBGWAS <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train,roc_cytokine_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train,roc_cytokine_test,roc_cytokine_test_HIVpos,roc_cytokine_test_HIVneg)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_cytokine_train,roc_cytokine_test,roc_cytokine_test_noTBSA,roc_cytokine_test_noTBGWAS)
```


### Selected metabolites in the 5-marker signature

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_metabols <- c("METHIONINE", "OXOPROLINE")
selected_metabols

data <-as.data.frame(metaboData.train[,selected_metabols,drop=FALSE])
data$Class = as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(metaboData.test[,selected_metabols,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(metaboData.test.HIVpos[,selected_metabols,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(metaboData.test.HIVneg[,selected_metabols,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(metaboData.test.noTBSA[,selected_metabols,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(metaboData.test.noTBGWAS[,selected_metabols,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```


#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_metabol_train <- predict(svmTune, data)
prob_metabol_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_metabol_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_metabol_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_metabol_test <- predict(svmTune, data_test)
prob_metabol_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_metabol_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_metabol_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_metabol_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVneg <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM model

```{r}
pred_metabol_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_metabol_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class, 
            predictor = prob_metabol_test_noTBSA[,1], 
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_noTBSA <- roc0
```










##### Entire test data (only no TB TBGWAS)
###### Test SVM svmTune 

```{r}
pred_metabol_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_metabol_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class, 
            predictor = prob_metabol_test_noTBGWAS[,1], 
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_noTBGWAS <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_metabol_train,roc_metabol_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_metabol_train,roc_metabol_test,roc_metabol_test_HIVpos,roc_metabol_test_HIVneg)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_metabol_train,roc_metabol_test,roc_metabol_test_noTBSA,roc_metabol_test_noTBGWAS)
```



### Combined model

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <- c("IFN.gamma" ,"IL.22"   ,  "IL.10"   )
selected_metabols <- selected_metabols <- c("METHIONINE", "OXOPROLINE")
selected_combined <- c("IFN.gamma" , "IL.22"   ,  "IL.10" ,"METHIONINE", "OXOPROLINE")

data <- as.data.frame(cbind(cytokineData.train[,selected_cytokines,drop=FALSE],metaboData.train[,selected_metabols,drop=FALSE]))
data$Class = as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cbind(cytokineData.test[,selected_cytokines,drop=FALSE],metaboData.test[,selected_metabols,drop=FALSE]))
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cbind(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE],metaboData.test.HIVpos[,selected_metabols,drop=FALSE]))
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cbind(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE],metaboData.test.HIVneg[,selected_metabols,drop=FALSE]))
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cbind(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE],metaboData.test.noTBSA[,selected_metabols,drop=FALSE]))
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cbind(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE],metaboData.test.noTBGWAS[,selected_metabols,drop=FALSE]))
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_combined_train <- predict(svmTune, data)
prob_combined_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_combined_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_combined_test <- predict(svmTune, data_test)
prob_combined_test <- predict(svmTune, data_test, type="prob")

ptb_prob <- prob_combined_test[, "Yes"]
cor.test(ptb_prob, sampleData.test$AGE, method="spearman")

df <- sampleData.test
df$Score <- ptb_prob
lm(Score ~ TB + AGE + Sex + HIV, data = df)
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_combined_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_combined_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_combined_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_combined_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_combined_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_combined_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_combined_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_combined_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_HIVneg <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_combined_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_combined_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_combined_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_combined_test_noTBSA <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_combined_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_combined_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_combined_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_noTBGWAS <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_combined_train,roc_combined_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_combined_train,roc_combined_test,roc_combined_test_HIVpos,roc_combined_test_HIVneg)
result <- threerocPlot_func_HIV(roc_combined_test,roc_combined_test_HIVpos,roc_combined_test_HIVneg)
result$plot
SuppFig5 <- result$plot
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_combined_train,roc_combined_test,roc_combined_test_noTBSA,roc_combined_test_noTBGWAS)
```



### Genarate figures
```{r}
result <- threerocPlot_func(roc_combined_test, roc_cytokine_test, roc_metabol_test)
result$plot  
Fig2_total <- result$plot 
result <- threerocPlot_func(roc_combined_test, roc_combined_test_HIVpos, roc_combined_test_HIVneg, names=c("Combined","HIV-positive","HIV-negative"),colors=c("#50b957", "red","#d89f32"))
result$plot  
Fig2_combined_HIV <- result$plot 
result <- threerocPlot_func(roc_combined_test, roc_combined_test_noTBGWAS, roc_combined_test_noTBSA,names=c("Combined","Symptomatic contacts","Hospitalized ORD"),colors=c("#50b957", "orange","#729cda"))
result$plot  
Fig2_combined_noTB <- result$plot 
result <- threerocPlot_func(roc_cytokine_test, roc_cytokine_test_HIVpos, roc_cytokine_test_HIVneg,names=c("Cytokine","HIV-positive","HIV-negative"),colors=c("#d84882",  "red","#d89f32"))
result$plot  
Fig2_cytokine_HIV <- result$plot 
result <- threerocPlot_func(roc_cytokine_test, roc_cytokine_test_noTBGWAS, roc_cytokine_test_noTBSA,names=c("Cytokine","Symptomatic contacts","Hospitalized ORD"),colors=c("#d84882","orange" ,"#729cda"))
result$plot  
Fig2_cytokine_noTB <- result$plot 
result <- threerocPlot_func(roc_metabol_test, roc_metabol_test_HIVpos, roc_metabol_test_HIVneg,
                            names=c("Metabolite","HIV-positive","HIV-negative"),colors=c("#7b63d4", "red","#d89f32"))
result$plot  
Fig2_metabol_HIV <- result$plot 
result <- threerocPlot_func(roc_metabol_test, roc_metabol_test_noTBGWAS, roc_metabol_test_noTBSA,names=c("Metabolite","Symptomatic contacts","Hospitalized ORD"),colors=c("#7b63d4", "orange","#729cda"))
result$plot  
Fig2_metabol_noTB <- result$plot 

save(Fig2_metabol_noTB,Fig2_metabol_HIV,Fig2_cytokine_noTB,Fig2_cytokine_HIV,Fig2_combined_noTB,Fig2_combined_HIV,Fig2_total,file="../Results/Fig2.RData")
```

## Make Figure 2
```{r}
l <- load("../Results/Fig2.RData")

Fig2_total$layers <-Fig2_total$layers[!sapply(Fig2_total$layers, function(x) inherits(x$geom, "GeomVline"))]
tpp_boxes <- data.frame(
  xmin = c(0.7, 0.98),   # Specificity lower limits
  xmax = c(1.0, 1.0),   # Specificity upper limits
  ymin = c(0.9, 0.65),  # Sensitivity lower limits
  ymax = c(1.0, 1),   # Sensitivity upper limits
  label = c("Triage Test", "POC Diagnostic")
)
Fig2_total <- Fig2_total+ geom_rect(data = tpp_boxes,
                                    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = label),
                                    alpha = 0.3, inherit.aes = FALSE) +
  scale_fill_manual(values = c("Triage Test" = "green", "POC Diagnostic" = "red"))+
  guides(color = guide_legend(title = "TPP Boxes"))+
  labs(title = "ROC Curves with Target Product Profile (TPP) Thresholds")+
  theme(title = element_text(size=12))


pdf(file="../Results/Figures/Manuscript/Figure 2.pdf",width=32,height = 30)
p1 <- plot_grid( NULL, Fig2_total,rel_widths = c(1,1.5),ncol = 2,labels = c("a", "b"),label_size = 36,label_fontface = "bold")
p2 <- plot_grid(Fig2_combined_HIV,Fig2_cytokine_HIV,Fig2_metabol_HIV,Fig2_combined_noTB,Fig2_cytokine_noTB,Fig2_metabol_noTB,nrow=2,ncol=3,labels = c("c","d","e","f","g","h"),label_size = 36,label_fontface = "bold",label_y = 1.1)
plot_grid(p1,p2,nrow=2,ncol=1,rel_heights = c(1,2))
dev.off()
```


## Data at baseline (only adults) - other signatures
### Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()
MetaboData_Baseline_Adult$Ratio <- ((2^MetaboData_Baseline_Adult$KYNURENINE)/(2^MetaboData_Baseline_Adult$TRYPTOPHAN))+(2^MetaboData_Baseline_Adult$RETINOL)
MetaboData_Baseline_Adult$Retinol <- (2^MetaboData_Baseline_Adult$RETINOL)
```

### Make train and test data  


```{r message=FALSE, warning=FALSE}
ClinicalData_Baseline_Adult$Group <- cbind(paste0(ClinicalData_Baseline_Adult$TB,",",ClinicalData_Baseline_Adult$HIV))

set.seed(123) 
test_percentage <- 0.3 

test_indices <- createDataPartition(ClinicalData_Baseline_Adult$Group, p = test_percentage, list = FALSE)

sampleData_train <- ClinicalData_Baseline_Adult[-test_indices, ]
sampleData_test <- ClinicalData_Baseline_Adult[test_indices, ]


removed_samles_train <- which(!sampleData_train$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_train <- sampleData_train[-removed_samles_train,]
removed_samles_test <- which(!sampleData_test$ID%in%row.names(CytokineData_Baseline_Adult))
sampleData_test <- sampleData_test[-removed_samles_test,]

table(sampleData_train$TB,sampleData_train$HIV)
table(sampleData_test$TB,sampleData_test$HIV)

metaboData.train <- MetaboData_Baseline_Adult[sampleData_train$ID,]
cytokineData.train <- CytokineData_Baseline_Adult_imputed[sampleData_train$ID,]
sampleData.train <- sampleData_train
metaboData.test <- MetaboData_Baseline_Adult[sampleData_test$ID,]
cytokineData.test <- CytokineData_Baseline_Adult_imputed[sampleData_test$ID,]
sampleData.test <- sampleData_test


metaboData.train <- as.data.frame(metaboData.train)
cytokineData.train <- as.data.frame(cytokineData.train)

sampleData_test_HIVpos <- sampleData_test%>%
  filter(HIV=="Positive")
sampleData_test_HIVneg <- sampleData_test%>%
  filter(HIV=="Negative")
sampleData_test_noTBSA <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "SA"))
sampleData_test_noTBGWAS <- sampleData_test%>%
  filter(TB == "Yes" | (TB == "No" & Study == "TBGWAS"))

metaboData.test.HIVpos <- metaboData.test[sampleData_test_HIVpos$ID,]
cytokineData.test.HIVpos <- cytokineData.test[sampleData_test_HIVpos$ID,]
sampleData.test.HIVpos <- sampleData_test_HIVpos

metaboData.test.HIVneg <- metaboData.test[sampleData_test_HIVneg$ID,]
cytokineData.test.HIVneg <- cytokineData.test[sampleData_test_HIVneg$ID,]
sampleData.test.HIVneg <- sampleData_test_HIVneg

metaboData.test.noTBSA <- metaboData.test[sampleData_test_noTBSA$ID,]
cytokineData.test.noTBSA <- cytokineData.test[sampleData_test_noTBSA$ID,]
sampleData.test.noTBSA <- sampleData_test_noTBSA

metaboData.test.noTBGWAS <- metaboData.test[sampleData_test_noTBGWAS$ID,]
cytokineData.test.noTBGWAS <- cytokineData.test[sampleData_test_noTBGWAS$ID,]
sampleData.test.noTBGWAS <- sampleData_test_noTBGWAS
```

### Selected metabolites (Ratio)

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
# selected_combined <- read_rds(file="../Results/randomforestSRC/selectedCombined_top7_baseline_adult.rds")
selected_metabols <- c("Ratio")
selected_metabols

data <-as.data.frame(metaboData.train[,selected_metabols,drop=FALSE])
data$Class = as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(metaboData.test[,selected_metabols,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(metaboData.test.HIVpos[,selected_metabols,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(metaboData.test.HIVneg[,selected_metabols,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(metaboData.test.noTBSA[,selected_metabols,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(metaboData.test.noTBGWAS[,selected_metabols,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```


#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_metabol_train <- predict(svmTune, data)
prob_metabol_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_metabol_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_metabol_train_ratio <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_metabol_test <- predict(svmTune, data_test)
prob_metabol_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_metabol_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_ratio <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_metabol_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVpos_ratio <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_metabol_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVneg_ratio <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM model

```{r}
pred_metabol_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_metabol_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class, 
            predictor = prob_metabol_test_noTBSA[,1], 
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_noTBSA_ratio <- roc0
```










##### Entire test data (only no TB TBGWAS)
###### Test SVM svmTune 

```{r}
pred_metabol_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_metabol_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class, 
            predictor = prob_metabol_test_noTBGWAS[,1], 
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_noTBGWAS_ratio <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_metabol_train_ratio,roc_metabol_test_ratio)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_metabol_train_ratio,roc_metabol_test_ratio,roc_metabol_test_HIVpos_ratio,roc_metabol_test_HIVneg_ratio)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_metabol_train_ratio,roc_metabol_test_ratio,roc_metabol_test_noTBSA_ratio,roc_metabol_test_noTBGWAS_ratio)
```

### Genarate figures
```{r}
result <- threerocPlot_func(roc_combined_test,roc_metabol_test,roc_metabol_test_ratio,
                            names=c("Combined","Metabolite","Kyn/Trp + retinol"),colors=c("#50b957","#7b63d4", "#3abbcc"))
result$plot  
Fig3_metabol <- result$plot 

result <- threerocPlot_func(roc_metabol_test_ratio, roc_metabol_test_HIVpos_ratio, roc_metabol_test_HIVneg_ratio,
                            names=c("Kyn/Trp + retinol","HIV-positive","HIV-negative"),colors=c("#3abbcc", "red","#d89f32"))
result$plot  
Fig3_metabol_HIV <- result$plot 
result <- threerocPlot_func(roc_metabol_test_ratio, roc_metabol_test_noTBGWAS_ratio, roc_metabol_test_noTBSA_ratio,names=c("Kyn/Trp + retinol","Symptomatic contacts","Hospitalized ORD"),colors=c("#3abbcc", "orange","#729cda"))
result$plot  
Fig3_metabol_noTB <- result$plot 

save(Fig3_metabol_noTB,Fig3_metabol_HIV,Fig3_metabol,file="../Results/Fig3_metabol.RData")
```


### Selected cytokines (IFNG)

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("IFN.gamma")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train_IFNG <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test_IFNG <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos_IFNG <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg_IFNG <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_cytokine_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_cytokine_test_noTBSA_IFNG <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_cytokine_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_noTBGWAS_IFNG <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train_IFNG,roc_cytokine_test_IFNG)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train_IFNG,roc_cytokine_test_IFNG,roc_cytokine_test_HIVpos_IFNG,roc_cytokine_test_HIVneg_IFNG)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_cytokine_train_IFNG,roc_cytokine_test_IFNG,roc_cytokine_test_noTBSA_IFNG,roc_cytokine_test_noTBGWAS_IFNG)
```

### Genarate figures
```{r}
result <- threerocPlot_func(roc_combined_test,roc_cytokine_test,roc_cytokine_test_IFNG,
                            names=c("Combined","Cytokine","IFN-gamma"),colors=c("#50b957","#c65999","#e08181"))
result$plot  
Fig3_cytokine <- result$plot 

result <- threerocPlot_func(roc_cytokine_test_IFNG, roc_cytokine_test_HIVpos_IFNG, roc_cytokine_test_HIVneg_IFNG,
                            names=c("IFN-gamma","HIV-positive","HIV-negative"),colors=c("#e08181", "red","#d89f32"))
result$plot  
Fig3_cytokine_HIV <- result$plot 
result <- threerocPlot_func(roc_cytokine_test_IFNG, roc_cytokine_test_noTBGWAS_IFNG, roc_cytokine_test_noTBSA_IFNG,names=c("IFN-gamma","Symptomatic contacts","Hospitalized ORD"),colors=c("#e08181", "orange","#729cda"))
result$plot  
Fig3_cytokine_noTB <- result$plot 

save(Fig3_cytokine_noTB,Fig3_cytokine_HIV,Fig3_cytokine,file="../Results/Fig3_cytokine.RData")
```

### Selected cytokines (IP.10)

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("IP.10")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train_IP10 <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test_IP10 <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos_IP10 <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg_IP10 <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_cytokine_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_cytokine_test_noTBSA_IP10 <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_cytokine_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_noTBGWAS_IP10 <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train_IP10,roc_cytokine_test_IP10)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train_IP10,roc_cytokine_test_IP10,roc_cytokine_test_HIVpos_IP10,roc_cytokine_test_HIVneg_IP10)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_cytokine_train_IP10,roc_cytokine_test_IP10,roc_cytokine_test_noTBSA_IP10,roc_cytokine_test_noTBGWAS_IP10)
```

### Selected cytokines (TNF.alpha)

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("TNF.alpha")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train_TNF <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test_TNF <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos_TNF <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg_TNF <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_cytokine_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_cytokine_test_noTBSA_TNF <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_cytokine_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_noTBGWAS_TNF <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train_TNF,roc_cytokine_test_TNF)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train_TNF,roc_cytokine_test_TNF,roc_cytokine_test_HIVpos_TNF,roc_cytokine_test_HIVneg_TNF)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_cytokine_train_TNF,roc_cytokine_test_TNF,roc_cytokine_test_noTBSA_TNF,roc_cytokine_test_noTBGWAS_TNF)
```

### Selected cytokines (IL6+IP.10+IL10)

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("IP.10","IL.6","IL.10")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))

data_test_noTBSA <- as.data.frame(cytokineData.test.noTBSA[,selected_cytokines,drop=FALSE])
data_test_noTBSA$Class = as.factor(gsub(" ","",sampleData.test.noTBSA$TB))
data_test_noTBSA$Class <- factor(data_test_noTBSA$Class, levels =c("Yes","No"))

data_test_noTBGWAS <- as.data.frame(cytokineData.test.noTBGWAS[,selected_cytokines,drop=FALSE])
data_test_noTBGWAS$Class = as.factor(gsub(" ","",sampleData.test.noTBGWAS$TB))
data_test_noTBGWAS$Class <- factor(data_test_noTBGWAS$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train_third <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test_third <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos_third <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg_third <- roc0
```










##### Entire test data (only no TB SA)
###### Test SVM

```{r}
pred_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA)
prob_cytokine_test_noTBSA <- predict(svmTune, data_test_noTBSA, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBSA$Class,
            predictor = prob_cytokine_test_noTBSA[,1],
            levels = rev(levels(data_test_noTBSA$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])
roc_cytokine_test_noTBSA_third <- roc0
```



##### Entire test data (only no TB TBGWAS)
###### Test SVM

```{r}
pred_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS)
prob_cytokine_test_noTBGWAS <- predict(svmTune, data_test_noTBGWAS, type="prob")
```


###### ROC curve

```{r}
roc0 <- roc(response = data_test_noTBGWAS$Class,
            predictor = prob_cytokine_test_noTBGWAS[,1],
            levels = rev(levels(data_test_noTBGWAS$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_noTBGWAS_third <- roc0
```



##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train_third,roc_cytokine_test_third)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train_third,roc_cytokine_test_third,roc_cytokine_test_HIVpos_third,roc_cytokine_test_HIVneg_third)
```

##### Both train and test and both no TB from SA and TBGWAS

```{r}
fourrocPlot_func_2(roc_cytokine_train_third,roc_cytokine_test_third,roc_cytokine_test_noTBSA_third,roc_cytokine_test_noTBGWAS_third)
```


### Selected cytokines (CRP)

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
selected_cytokines <-c("CRP")
selected_cytokines

CRP <- read_excel("../../Data/ReadyData/2025-03-03_PotPrev_Emory_Data.xlsx")
CRP <- CRP%>%filter(CRP=="Yes")
CRP$study_id <- paste0("SA",CRP$study_id)
CRP <- as.data.frame(CRP)
row.names(CRP) <- CRP$study_id
CRP <- CRP[,c("study_id","Results")]
names(CRP) <- c("study_id","CRP")

cytokineData.train <- CRP[intersect(row.names(cytokineData.train),CRP$study_id),"CRP",drop=FALSE]
row.names(sampleData.train) <- sampleData.train$ID
sampleData.train <- sampleData.train[row.names(cytokineData.train),]

cytokineData.test <- CRP[intersect(row.names(cytokineData.test),CRP$study_id),"CRP",drop=FALSE]
row.names(sampleData.test) <- sampleData.test$ID
sampleData.test <- sampleData.test[row.names(cytokineData.test),]

cytokineData.test.HIVpos <- CRP[intersect(row.names(cytokineData.test.HIVpos),CRP$study_id),"CRP",drop=FALSE]
row.names(sampleData.test.HIVpos) <- sampleData.test.HIVpos$ID
sampleData.test.HIVpos <- sampleData.test.HIVpos[row.names(cytokineData.test.HIVpos),]

cytokineData.test.HIVneg <- CRP[intersect(row.names(cytokineData.test.HIVneg),CRP$study_id),"CRP",drop=FALSE]
row.names(sampleData.test.HIVneg) <- sampleData.test.HIVneg$ID
sampleData.test.HIVneg <- sampleData.test.HIVneg[row.names(cytokineData.test.HIVneg),]

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$TB))
data$Class <- factor(data$Class, levels =c("Yes","No"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$TB))
data_test$Class <- factor(data_test$Class, levels =c("Yes","No"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$TB))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("Yes","No"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$TB))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("Yes","No"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train_CRP <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test_CRP <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos_CRP <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg_CRP <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train_CRP,roc_cytokine_test_CRP)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train_CRP,roc_cytokine_test_CRP,roc_cytokine_test_HIVpos_CRP,roc_cytokine_test_HIVneg_CRP)
```


### Genarate figures

```{r}
result <- threerocPlot_func(roc_cytokine_test_IP10,roc_cytokine_test_TNF,roc_cytokine_test_third,
                            names=c("IP.10","TNF.alpha","IP.10,IL.6,IL10"),colors=c("#6a732a","#c57ebf","#ae7d45" ))
result$plot  
Fig3_cytokine_others <- result$plot 

result <- threerocPlot_func(roc_cytokine_test_IP10, roc_cytokine_test_HIVpos_IP10, roc_cytokine_test_HIVneg_IP10,
                            names=c("IP.10","HIV-positive","HIV-negative"),colors=c("#6a732a", "red","#d89f32"))
result$plot  
Fig3_cytokine_HIV_IP10 <- result$plot 
result <- threerocPlot_func(roc_cytokine_test_IP10, roc_cytokine_test_noTBGWAS_IP10, roc_cytokine_test_noTBSA_IP10,names=c("IP.10","Symptomatic contacts","Hospitalized ORD"),colors=c("#6a732a", "orange","#729cda"))
result$plot  
Fig3_cytokine_noTB_IP10 <- result$plot 

result <- threerocPlot_func(roc_cytokine_test_TNF, roc_cytokine_test_HIVpos_TNF, roc_cytokine_test_HIVneg_TNF,
                            names=c("TNF.alpha","HIV-positive","HIV-negative"),colors=c("#c57ebf", "red","#d89f32"))
result$plot  
Fig3_cytokine_HIV_TNF <- result$plot 
result <- threerocPlot_func(roc_cytokine_test_TNF, roc_cytokine_test_noTBGWAS_TNF, roc_cytokine_test_noTBSA_TNF,names=c("TNF.alpha","Symptomatic contacts","Hospitalized ORD"),colors=c("#c57ebf", "orange","#729cda"))
result$plot  
Fig3_cytokine_noTB_TNF <- result$plot 

result <- threerocPlot_func(roc_cytokine_test_third, roc_cytokine_test_HIVpos_third, roc_cytokine_test_HIVneg_third,
                            names=c("IP.10,IL.6,IL10","HIV-positive","HIV-negative"),colors=c("#ae7d45", "red","#d89f32"))
result$plot  
Fig3_cytokine_HIV_third <- result$plot 
result <- threerocPlot_func(roc_cytokine_test_third, roc_cytokine_test_noTBGWAS_third, roc_cytokine_test_noTBSA_third,names=c("IP.10,IL.6,IL10","Symptomatic contacts","Hospitalized ORD"),colors=c("#ae7d45", "orange","#729cda"))
result$plot  
Fig3_cytokine_noTB_third <- result$plot 

save(Fig3_cytokine_noTB_IP10,Fig3_cytokine_noTB_TNF,Fig3_cytokine_noTB_third,Fig3_cytokine_HIV_IP10,Fig3_cytokine_HIV_TNF,Fig3_cytokine_HIV_third,Fig3_cytokine_others,file="../Results/Fig3_cytokine_others.RData")
```

## Make Figure 3

```{r}
pdf(file="../Results/Figures/Manuscript/Figure3.pdf",width=32,height = 20)
plot_grid(Fig3_metabol,Fig3_metabol_HIV,Fig3_metabol_noTB,ncol=3,labels=c("a","b","c"),label_size = 36)/plot_grid(Fig3_cytokine,Fig3_cytokine_HIV,Fig3_cytokine_noTB,ncol=3,labels = c("d","e","f"),label_size = 36)
dev.off()
```

## Make Supplementary Figure 4

```{r}
pdf(file="../Results/Figures/Manuscript/SuppFig4.pdf",width=40,height = 20)
p1 <- plot_grid(Fig3_cytokine_HIV_IP10,Fig3_cytokine_HIV_TNF,Fig3_cytokine_HIV_third,Fig3_cytokine_noTB_IP10,Fig3_cytokine_noTB_TNF,Fig3_cytokine_noTB_third,ncol=3,nrow=2,labels = c("b","c","d","e","f","g"),label_size = 36)
p2 <- plot_grid(Fig3_cytokine_others,labels = c("a"),label_size = 36)
plot_grid(p2,p1,rel_heights = c(1,1),rel_widths = c(0.5,1.5))
dev.off()
```



## Data at baseline (only adults) - DE Metabolites for treatment response

### Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Load data

```{r}
MetaboData_Baseline_Adult <- MetaboData_Baseline_Adult[ClinicalData_Baseline_Adult$ID,]
dim(MetaboData_Baseline_Adult)
```


### Creating a covariates matrix

```{r}
MetaboData_Baseline_Adult <- MetaboData_Baseline_Adult%>%t()
covareiateData <- ClinicalData_Baseline_Adult[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult$Outcome)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("Cure","Failure","Age","HIV","Sex")
cont.matrix <- makeContrasts(Cure-Failure,levels=design)
fit <- lmFit(MetaboData_Baseline_Adult, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "fdr"))
volcanoPlot_data_MET <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_MET) <- c("logFC.Cure.Failure","pValue.Cure.Failure","qValue.Cure.Failure")

volcanoPlot_data_MET$Metabolite <- row.names(volcanoPlot_data_MET)
save(volcanoPlot_data_MET,file="../Results/DEA/DEM_Baseline_Adult_Treatment.RData")
DEGs <- c(volcanoPlot_data_MET%>%
            filter(logFC.Cure.Failure>0&pValue.Cure.Failure<0.05)%>%row.names(),
          volcanoPlot_data_MET%>%
            filter(logFC.Cure.Failure<0&pValue.Cure.Failure<0.05)%>%row.names())

volcanoPlot_data_MET%>%
  filter(logFC.Cure.Failure>0&qValue.Cure.Failure<0.05)%>%
  nrow()
volcanoPlot_data_MET%>%
  filter(logFC.Cure.Failure<0&qValue.Cure.Failure<0.05)%>%
  nrow()
```

### Volcano plot for Cure vs. Failure

#### Using p-value 

```{r message=FALSE, warning=FALSE}
Fig1_a_volcano <- volcanoPlot_func_2(volcanoPlot_data_MET, 
                                     xaxis = "logFC.Cure.Failure", 
                                     yaxis = "pValue.Cure.Failure", 
                                     uplab = "Up in PTB-Cured", 
                                     downlab = "Down in PTB-Cured", 
                                     name = "Metabolite",
                                     thr = 0.05, thr2 = 0, 
                                     label_points = FALSE,
                                     xlim_range = c(-3, 3),
                                     highlight_names = c("VALINE","MALATE",DEGs))+
  theme(
    legend.position = c(0.025, 0.025),
    legend.justification = c(0, 0)
  )
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func(volcanoPlot_data_MET, 
                 xaxis = "logFC.Cure.Failure", 
                 yaxis = "qValue.Cure.Failure", 
                 uplab = "Up in PTB-Cured", 
                 downlab = "Down in PTB-Cured", 
                 name = "Metabolite",
                 thr = 0.05, thr2 = 0, 
                 label_points = FALSE,
                 xlim_range = c(-3, 3),
                 highlight_names = c("VALINE","MALATE",DEGs))+
  theme(legend.position = "none")
```

## Data at baseline (only adults) - DE Cytokines for treatment response

### Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Load data

```{r}
CytokineData_Baseline_Adult_imputed <- CytokineData_Baseline_Adult_imputed[ClinicalData_Baseline_Adult$ID,]
dim(CytokineData_Baseline_Adult_imputed)
```


### Creating a covariates matrix

```{r}
CytokineData_Baseline_Adult_imputed <- CytokineData_Baseline_Adult_imputed%>%t()
covareiateData <- ClinicalData_Baseline_Adult[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult$Outcome)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("Cure","Failure","Age","HIV","Sex")
cont.matrix <- makeContrasts(Cure-Failure,levels=design)
fit <- lmFit(CytokineData_Baseline_Adult_imputed, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "BH"))
volcanoPlot_data_CYT <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_CYT) <- c("logFC.Cure.Failure","pValue.Cure.Failure","qValue.Cure.Failure")

volcanoPlot_data_CYT$Cytokine <- row.names(volcanoPlot_data_CYT)
save(volcanoPlot_data_CYT,file="../Results/DEA/DEC_Baseline_Adult_Treatment.RData")
DEGs <- c(volcanoPlot_data_CYT%>%
            filter(logFC.Cure.Failure>0&pValue.Cure.Failure<0.05)%>%row.names(),
          volcanoPlot_data_CYT%>%
            filter(logFC.Cure.Failure<0&pValue.Cure.Failure<0.05)%>%row.names())
volcanoPlot_data_CYT%>%
  filter(logFC.Cure.Failure>0&qValue.Cure.Failure<0.05)%>%
  nrow()
volcanoPlot_data_CYT%>%
  filter(logFC.Cure.Failure<0&qValue.Cure.Failure<0.05)%>%
  nrow()
```

### Volcano plot for Cure vs. Failure

#### Using p-value 

```{r message=FALSE, warning=FALSE}
Fig1_b_volcano <- volcanoPlot_func_2(volcanoPlot_data_CYT, 
                                     xaxis = "logFC.Cure.Failure", 
                                     yaxis = "pValue.Cure.Failure", 
                                     uplab = "Up in PTB-Cured", 
                                     downlab = "Down in PTB-Cured", 
                                     name = "Cytokine",
                                     thr = 0.05, thr2 = 0, 
                                     label_points = FALSE,
                                     xlim_range = c(-4, 4),
                                     highlight_names = c("IL.17A",DEGs))+
  theme(legend.position = "none")
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func(volcanoPlot_data_CYT, 
                 xaxis = "logFC.Cure.Failure", 
                 yaxis = "qValue.Cure.Failure", 
                 uplab = "Up in PTB-Cured", 
                 downlab = "Down in PTB-Cured", 
                 name = "Cytokine",
                 thr = 0.05, thr2 = 0, 
                 label_points = FALSE,
                 xlim_range = c(-4, 4),
                 highlight_names = c("IL.17A"))
```


## Data at Week 8 (only adults) - DE Metabolites for treatment response

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
SA_samples$LONGID <- gsub("SA","",SA_samples$ID)
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))

ClinicalData_Baseline_Adult_TreatmentFollowUp[match(SA_samples$ID,ClinicalData_Baseline_Adult_TreatmentFollowUp$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Visit.Week==8,TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Load data

```{r}
MetaboData_Baseline_Adult_TreatmentFollowUp <- MetaboData_Baseline_Adult_TreatmentFollowUp[ClinicalData_Baseline_Adult_TreatmentFollowUp$ID,]
dim(MetaboData_Baseline_Adult_TreatmentFollowUp)
```


### Creating a covariates matrix

```{r}
MetaboData_Baseline_Adult_TreatmentFollowUp <- MetaboData_Baseline_Adult_TreatmentFollowUp%>%t()
covareiateData <- ClinicalData_Baseline_Adult_TreatmentFollowUp[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Outcome)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("Cure","Failure","Age","HIV","Sex")
cont.matrix <- makeContrasts(Cure-Failure,levels=design)
fit <- lmFit(MetaboData_Baseline_Adult_TreatmentFollowUp, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "fdr"))
volcanoPlot_data_MET <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_MET) <- c("logFC.Cure.Failure","pValue.Cure.Failure","qValue.Cure.Failure")

volcanoPlot_data_MET$Metabolite <- row.names(volcanoPlot_data_MET)
save(volcanoPlot_data_MET,file="../Results/DEA/DEM_week8_Adult_Treatment.RData")
DEGs <- c(volcanoPlot_data_MET%>%
            filter(logFC.Cure.Failure>0&pValue.Cure.Failure<0.05)%>%row.names(),
          volcanoPlot_data_MET%>%
            filter(logFC.Cure.Failure<0&pValue.Cure.Failure<0.05)%>%row.names())
volcanoPlot_data_MET%>%
  filter(logFC.Cure.Failure>0&qValue.Cure.Failure<0.05)%>%
  nrow()
volcanoPlot_data_MET%>%
  filter(logFC.Cure.Failure<0&qValue.Cure.Failure<0.05)%>%
  nrow()
```

### Volcano plot for Cure vs. Failure

#### Using p-value 

```{r message=FALSE, warning=FALSE}
Fig2_a_volcano <- volcanoPlot_func_2(volcanoPlot_data_MET, 
                                     xaxis = "logFC.Cure.Failure", 
                                     yaxis = "pValue.Cure.Failure", 
                                     uplab = "Up in PTB-Cured", 
                                     downlab = "Down in PTB-Cured", 
                                     name = "Metabolite",
                                     thr = 0.05, thr2 = 0, 
                                     label_points = FALSE,
                                     xlim_range = c(-3, 3),
                                     highlight_names = c(DEGs))+
  theme(legend.position = "none")
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func(volcanoPlot_data_MET, 
                 xaxis = "logFC.Cure.Failure", 
                 yaxis = "qValue.Cure.Failure", 
                 uplab = "Up in PTB-Cured", 
                 downlab = "Down in PTB-Cured", 
                 name = "Metabolite",
                 thr = 0.05, thr2 = 0, 
                 label_points = FALSE,
                 xlim_range = c(-3, 3),
                 highlight_names = c("URATE","ANSERINE"))
```

## Data at Week 8 (only adults) - DE Cytokines for treatment response

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
SA_samples$LONGID <- gsub("SA","",SA_samples$ID)
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))

ClinicalData_Baseline_Adult_TreatmentFollowUp[match(SA_samples$ID,ClinicalData_Baseline_Adult_TreatmentFollowUp$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Visit.Week==8,TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Load data

```{r}
CytokineData_Baseline_Adult_TreatmentFollowUp_imputed <- CytokineData_Baseline_Adult_TreatmentFollowUp_imputed[ClinicalData_Baseline_Adult_TreatmentFollowUp$ID,]
dim(CytokineData_Baseline_Adult_TreatmentFollowUp_imputed)
```


### Creating a covariates matrix

```{r}
CytokineData_Baseline_Adult_TreatmentFollowUp_imputed <- CytokineData_Baseline_Adult_TreatmentFollowUp_imputed%>%t()
covareiateData <- ClinicalData_Baseline_Adult_TreatmentFollowUp[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Outcome)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("Cure","Failure","Age","HIV","Sex")
cont.matrix <- makeContrasts(Cure-Failure,levels=design)
fit <- lmFit(CytokineData_Baseline_Adult_TreatmentFollowUp_imputed, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "BH"))
volcanoPlot_data_CYT <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_CYT) <- c("logFC.Cure.Failure","pValue.Cure.Failure","qValue.Cure.Failure")

volcanoPlot_data_CYT$Cytokine <- row.names(volcanoPlot_data_CYT)
save(volcanoPlot_data_CYT,file="../Results/DEA/DEC_week8_Adult_Treatment.RData")
DEGs <- c(volcanoPlot_data_CYT%>%
            filter(logFC.Cure.Failure>0&pValue.Cure.Failure<0.05)%>%row.names(),
          volcanoPlot_data_CYT%>%
            filter(logFC.Cure.Failure<0&pValue.Cure.Failure<0.05)%>%row.names())

volcanoPlot_data_CYT%>%
  filter(logFC.Cure.Failure>0&qValue.Cure.Failure<0.05)%>%
  nrow()
volcanoPlot_data_CYT%>%
  filter(logFC.Cure.Failure<0&qValue.Cure.Failure<0.05)%>%
  nrow()
```

### Volcano plot for Cure vs. Failure

#### Using p-value 

```{r message=FALSE, warning=FALSE}
Fig2_b_volcano <- volcanoPlot_func_2(volcanoPlot_data_CYT, 
                                     xaxis = "logFC.Cure.Failure", 
                                     yaxis = "pValue.Cure.Failure", 
                                     uplab = "Up in PTB-Cured", 
                                     downlab = "Down in PTB-Cured", 
                                     name = "Cytokine",
                                     thr = 0.05, thr2 = 0, 
                                     label_points = FALSE,
                                     xlim_range = c(-4, 4),
                                     highlight_names = c(DEGs))+
  theme(legend.position = "none")
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func(volcanoPlot_data_CYT, 
                 xaxis = "logFC.Cure.Failure", 
                 yaxis = "qValue.Cure.Failure", 
                 uplab = "Up in PTB-Cured", 
                 downlab = "Down in PTB-Cured", 
                 name = "Cytokine",
                 thr = 0.05, thr2 = 0, 
                 label_points = FALSE,
                 xlim_range = c(-4, 4),
                 highlight_names = c("IL.8"))
```



## Data at Week 17 or 26 (only adults) - DE Metabolites for treatment response

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
SA_samples$LONGID <- gsub("SA","",SA_samples$ID)
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))

ClinicalData_Baseline_Adult_TreatmentFollowUp[match(SA_samples$ID,ClinicalData_Baseline_Adult_TreatmentFollowUp$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Visit.Week==17|Visit.Week==26)
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Load data

```{r}
MetaboData_Baseline_Adult_TreatmentFollowUp <- MetaboData_Baseline_Adult_TreatmentFollowUp[ClinicalData_Baseline_Adult_TreatmentFollowUp$ID,]
dim(MetaboData_Baseline_Adult_TreatmentFollowUp)
```


### Creating a covariates matrix

```{r}
MetaboData_Baseline_Adult_TreatmentFollowUp <- MetaboData_Baseline_Adult_TreatmentFollowUp%>%t()
covareiateData <- ClinicalData_Baseline_Adult_TreatmentFollowUp[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Outcome)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("Cure","Failure","Age","HIV","Sex")
cont.matrix <- makeContrasts(Cure-Failure,levels=design)
fit <- lmFit(MetaboData_Baseline_Adult_TreatmentFollowUp, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "fdr"))
volcanoPlot_data_MET <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_MET) <- c("logFC.Cure.Failure","pValue.Cure.Failure","qValue.Cure.Failure")

volcanoPlot_data_MET$Metabolite <- row.names(volcanoPlot_data_MET)
save(volcanoPlot_data_MET,file="../Results/DEA/DEM_week17or26_Adult_Treatment.RData")
DEGs <- c(volcanoPlot_data_MET%>%
            filter(logFC.Cure.Failure>0&pValue.Cure.Failure<0.05)%>%row.names(),
          volcanoPlot_data_MET%>%
            filter(logFC.Cure.Failure<0&pValue.Cure.Failure<0.05)%>%row.names())

volcanoPlot_data_MET%>%
  filter(logFC.Cure.Failure>0&qValue.Cure.Failure<0.05)%>%
  nrow()
volcanoPlot_data_MET%>%
  filter(logFC.Cure.Failure<0&qValue.Cure.Failure<0.05)%>%
  nrow()
```

### Volcano plot for Cure vs. Failure

#### Using p-value 

```{r message=FALSE, warning=FALSE}
Fig3_a_volcano <- volcanoPlot_func_2(volcanoPlot_data_MET, 
                                     xaxis = "logFC.Cure.Failure", 
                                     yaxis = "pValue.Cure.Failure", 
                                     uplab = "Up in PTB-Cured", 
                                     downlab = "Down in PTB-Cured", 
                                     name = "Metabolite",
                                     thr = 0.05, thr2 = 0, 
                                     label_points = FALSE,
                                     xlim_range = c(-3, 3),
                                     highlight_names = c(DEGs))+
  theme(legend.position = "none")
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func(volcanoPlot_data_MET, 
                 xaxis = "logFC.Cure.Failure", 
                 yaxis = "qValue.Cure.Failure", 
                 uplab = "Up in PTB-Cured", 
                 downlab = "Down in PTB-Cured", 
                 name = "Metabolite",
                 thr = 0.05, thr2 = 0, 
                 label_points = FALSE,
                 xlim_range = c(-3, 3),
                 highlight_names = c(DEGs))+
  theme(legend.position = "none")
```

## Data at Week 17 or 26 (only adults) - DE Cytokines for treatment response

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
SA_samples$LONGID <- gsub("SA","",SA_samples$ID)
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))

ClinicalData_Baseline_Adult_TreatmentFollowUp[match(SA_samples$ID,ClinicalData_Baseline_Adult_TreatmentFollowUp$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Visit.Week==17|Visit.Week==26)
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Load data

```{r}
CytokineData_Baseline_Adult_TreatmentFollowUp_imputed <- CytokineData_Baseline_Adult_TreatmentFollowUp_imputed[ClinicalData_Baseline_Adult_TreatmentFollowUp$ID,]
dim(CytokineData_Baseline_Adult_TreatmentFollowUp_imputed)
```


### Creating a covariates matrix

```{r}
CytokineData_Baseline_Adult_TreatmentFollowUp_imputed <- CytokineData_Baseline_Adult_TreatmentFollowUp_imputed%>%t()
covareiateData <- ClinicalData_Baseline_Adult_TreatmentFollowUp[,c("AGE","HIV","Sex"),drop=FALSE]
```

### Model

```{r message=FALSE, warning=FALSE}
f <- factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Outcome)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("Cure","Failure","Age","HIV","Sex")
cont.matrix <- makeContrasts(Cure-Failure,levels=design)
fit <- lmFit(CytokineData_Baseline_Adult_TreatmentFollowUp_imputed, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)
```

### Visualize Limma results for metabolite data

```{r message=FALSE, warning=FALSE}
logFC <- as.data.frame(fit.b$coefficients)
pValue <- as.data.frame(fit.b$p.value)
qValue <- apply(pValue,2,function(x) p.adjust(x,method = "BH"))
volcanoPlot_data_CYT <- cbind(logFC=logFC,pValue=pValue,qValue=qValue)
names(volcanoPlot_data_CYT) <- c("logFC.Cure.Failure","pValue.Cure.Failure","qValue.Cure.Failure")

volcanoPlot_data_CYT$Cytokine <- row.names(volcanoPlot_data_CYT)
save(volcanoPlot_data_CYT,file="../Results/DEA/DEC_week17or26_Adult_Treatment.RData")
DEGs <- c(volcanoPlot_data_CYT%>%
            filter(logFC.Cure.Failure>0&pValue.Cure.Failure<0.05)%>%row.names(),
          volcanoPlot_data_CYT%>%
            filter(logFC.Cure.Failure<0&pValue.Cure.Failure<0.05)%>%row.names())
volcanoPlot_data_CYT%>%
  filter(logFC.Cure.Failure>0&qValue.Cure.Failure<0.05)%>%
  nrow()
volcanoPlot_data_CYT%>%
  filter(logFC.Cure.Failure<0&qValue.Cure.Failure<0.05)%>%
  nrow()
```

### Volcano plot for Cure vs. Failure

#### Using p-value 

```{r message=FALSE, warning=FALSE}
Fig3_b_volcano <- volcanoPlot_func_2(volcanoPlot_data_CYT, 
                                     xaxis = "logFC.Cure.Failure", 
                                     yaxis = "pValue.Cure.Failure", 
                                     uplab = "Up in PTB-Cured", 
                                     downlab = "Down in PTB-Cured", 
                                     name = "Cytokine",
                                     thr = 0.05, thr2 = 0, 
                                     label_points = FALSE,
                                     xlim_range = c(-4, 4),
                                     highlight_names = c(DEGs))+
  theme(legend.position = "none")
```

#### Using q-value 

```{r message=FALSE, warning=FALSE}
volcanoPlot_func(volcanoPlot_data_CYT, 
                 xaxis = "logFC.Cure.Failure", 
                 yaxis = "qValue.Cure.Failure", 
                 uplab = "Up in PTB-Cured", 
                 downlab = "Down in PTB-Cured", 
                 name = "Cytokine",
                 thr = 0.05, thr2 = 0, 
                 label_points = FALSE,
                 xlim_range = c(-4, 4),
                 highlight_names = c(DEGs))+
  theme(legend.position = "none")
```


## Make Supplementary Figure 5

```{r}
Fig <- plot_grid(
  plot_grid(
    Fig1_a_volcano, Fig2_a_volcano, Fig3_a_volcano,
    labels = c("a", "c", "e"),
    label_size = 36,
    nrow = 3,
    rel_heights = c(1,1,1)
  ),
  plot_grid(
    Fig1_b_volcano, Fig2_b_volcano, Fig3_b_volcano,
    labels = c("b", "d", "f"),
    label_size = 36,
    nrow = 3,
    rel_heights = c(1,1,1)
  ),
  ncol = 2,
  rel_widths = c(1, 1)
)

pdf(file="../Results/Figures/Manuscript/SuppFig5.pdf",height = 30,width = 20)
Fig
dev.off()
```

## Data at baseline (only adults) - Multi-omics biomarker for predicting treatment response including Boruta
### Load metabolite, cytokine and clinical data 

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(TB=="Yes",Outcome%in%c("FAILED","CURED"))
```

### Make train and test data  

```{r message=FALSE, warning=FALSE}
ClinicalData_Baseline_Adult$Group <- cbind(paste0(ClinicalData_Baseline_Adult$Outcome,",",ClinicalData_Baseline_Adult$HIV))

set.seed(123) 
test_percentage <- 0.3 # specify the percentage of data to be used for testing

test_indices <- createDataPartition(ClinicalData_Baseline_Adult$Group, p = test_percentage, list = FALSE)

sampleData_train <- ClinicalData_Baseline_Adult[-test_indices, ]
sampleData_test <- ClinicalData_Baseline_Adult[test_indices, ]

# library(caret)
# 
# sampleData_train$Outcome <- as.factor(sampleData_train$Outcome)
# 
# 
# train_down <-downSample(x = sampleData_train[, !(names(sampleData_train) %in% "Outcome")],
#                      y = sampleData_train$Outcome)
# 
# # caret renames the outcome column to "Class"
# table(train_down$Class)
# names(train_down)[names(train_down) == "Class"] <- "Outcome"
# sampleData_train <- train_down

metaboData.train <- MetaboData_Baseline_Adult[sampleData_train$ID,]
cytokineData.train <- CytokineData_Baseline_Adult_imputed[sampleData_train$ID,]
sampleData.train <- sampleData_train
metaboData.test <- MetaboData_Baseline_Adult[sampleData_test$ID,]
cytokineData.test <- CytokineData_Baseline_Adult_imputed[sampleData_test$ID,]
sampleData.test <- sampleData_test

metaboData.train <- as.data.frame(metaboData.train)
cytokineData.train <- as.data.frame(cytokineData.train)

sampleData.test.HIVpos <- sampleData.test%>%
  filter(HIV=="Positive")
sampleData.test.HIVneg <- sampleData.test%>%
  filter(HIV=="Negative")


metaboData.test.HIVpos <- metaboData.test[sampleData.test.HIVpos$ID,]
cytokineData.test.HIVpos <- cytokineData.test[sampleData.test.HIVpos$ID,]
sampleData.test.HIVpos <- sampleData.test.HIVpos

metaboData.test.HIVneg <- metaboData.test[sampleData.test.HIVneg$ID,]
cytokineData.test.HIVneg <- cytokineData.test[sampleData.test.HIVneg$ID,]
sampleData.test.HIVneg <- sampleData.test.HIVneg


table(sampleData.train$TB,sampleData.train$Outcome)
table(sampleData.test$TB,sampleData.test$Outcome)
```

### Make a combined data including metabolites and cytokine for feature selection

```{r}
combinedData.train <- cbind(metaboData.train,cytokineData.train)
combinedData.test <- cbind(metaboData.test,cytokineData.test)
combinedData.test.HIVpos <- cbind(metaboData.test.HIVpos,cytokineData.test.HIVpos)
combinedData.test.HIVneg <- cbind(metaboData.test.HIVneg,cytokineData.test.HIVneg)
combinedData.train$Group= as.factor(sampleData.train$Outcome)
combinedData.train$Group <- factor(combinedData.train$Group, levels =c("CURED","FAILED"))
```

### Run randomforestSRC

```{r eval=FALSE}
set.seed(111)
rf_model <- rfsrc(Group ~ ., data = combinedData.train, ntree = 500, importance = TRUE)
vimp <- rf_model$importance
vimp[order(vimp[,1],decreasing = TRUE),]
vimp.rfsrc(rf_model)
min_depth <- var.select(rf_model)
selected_combined_SRC <- min_depth$md.obj$topvars
selected_combined_SRC_top2 <- head(min_depth$topvars, 2)
selected_combined_SRC_top5 <- head(min_depth$topvars, 5)
selected_combined_SRC_top7 <- head(min_depth$topvars, 7)
saveRDS(selected_combined_SRC,file="../Results/randomforestSRC/selectedCombined_baseline_adult.rds")
saveRDS(selected_combined_SRC_top2, file="../Results/randomforestSRC/selectedCombined_top2_baseline_adult.rds")
saveRDS(selected_combined_SRC_top5, file="../Results/randomforestSRC/selectedCombined_top5_baseline_adult.rds")
saveRDS(selected_combined_SRC_top7, file="../Results/randomforestSRC/selectedCombined_top7_baseline_adult.rds")
```

### Run Boruta

```{r eval=FALSE}
set.seed(111)
boruta_data <- Boruta(Group~., data = combinedData.train, doTrace = 2, maxRuns=400)
boruta_data <- TentativeRoughFix(boruta_data)
boruta_stats <- attStats(boruta_data)
boruta_stats$cytokine <- rownames(boruta_stats)
boruta_stats <- boruta_stats %>% filter(decision =="Confirmed")
boruta_stats <- boruta_stats[order(boruta_stats$meanImp,decreasing = TRUE),c("cytokine","meanImp")]
plot(boruta_data, cex.axis = 1.1, xlab = "", las = 2,ylab="")
title("Cytokine",line=+1,adj=0,cex.main=2)
combined_boruta <- boruta_stats$cytokine
combined_boruta
save(combined_boruta,file="../Results/Boruta/selected_metabols_treatment.RData")
load("../Results/Boruta/selected_metabols_treatment.RData")
```


### Selected cytokines from Boruta

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
# selected_combined <- read_rds(file="../Results/randomforestSRC/selectedCombined_top7_baseline_adult.rds")
selected_cytokines <-c("IL.17A")
selected_cytokines

data <- as.data.frame(cytokineData.train[,selected_cytokines ,drop=FALSE])
data$Class= as.factor(gsub(" ","",sampleData.train$Outcome))
data$Class <- factor(data$Class, levels =c("CURED","FAILED"))

data_test <- as.data.frame(cytokineData.test[,selected_cytokines,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$Outcome))
data_test$Class <- factor(data_test$Class, levels =c("CURED","FAILED"))

data_test_HIVpos <- as.data.frame(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$Outcome))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("CURED","FAILED"))

data_test_HIVneg <- as.data.frame(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$Outcome))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("CURED","FAILED"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_cytokine_train <- predict(svmTune, data)
prob_cytokine_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_cytokine_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_cytokine_test <- predict(svmTune, data_test)
prob_cytokine_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_cytokine_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_cytokine_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_cytokine_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_cytokine_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_cytokine_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_cytokine_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_cytokine_test_HIVneg <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_cytokine_train,roc_cytokine_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_cytokine_train,roc_cytokine_test,roc_cytokine_test_HIVpos,roc_cytokine_test_HIVneg)
```

### Selected metabolites from Boruta

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
# selected_combined <- read_rds(file="../Results/randomforestSRC/selectedCombined_top7_baseline_adult.rds")
selected_metabols <- c("TYROSINE","ARACHIDONIC ACID", "MALATE" ,"EICOSAPENTAENOIC ACID" ,"VALINE")
# selected_metabols <- c("TYROSINE","EICOSAPENTAENOIC ACID" ,"VALINE","ARACHIDONIC ACID")
# selected_metabols <- c("TYROSINE","EICOSAPENTAENOIC ACID" ,"ARACHIDONIC ACID")
# selected_metabols <- selected_metabols[1:2]

data <-as.data.frame(metaboData.train[,selected_metabols,drop=FALSE])
data$Class = as.factor(gsub(" ","",sampleData.train$Outcome))
data$Class <- factor(data$Class, levels =c("CURED","FAILED"))

data_test <- as.data.frame(metaboData.test[,selected_metabols,drop=FALSE])
data_test$Class = as.factor(gsub(" ","",sampleData.test$Outcome))
data_test$Class <- factor(data_test$Class, levels =c("CURED","FAILED"))

data_test_HIVpos <- as.data.frame(metaboData.test.HIVpos[,selected_metabols,drop=FALSE])
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$Outcome))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("CURED","FAILED"))

data_test_HIVneg <- as.data.frame(metaboData.test.HIVneg[,selected_metabols,drop=FALSE])
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$Outcome))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("CURED","FAILED"))

```


#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_metabol_train <- predict(svmTune, data)
prob_metabol_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_metabol_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_metabol_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_metabol_test <- predict(svmTune, data_test)
prob_metabol_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_metabol_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_metabol_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_metabol_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_metabol_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_metabol_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_metabol_test_HIVneg <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_metabol_train,roc_metabol_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_metabol_train,roc_metabol_test,roc_metabol_test_HIVpos,roc_metabol_test_HIVneg)
```

### Combined model

#### Make data ready for SVM

```{r message=FALSE, warning=FALSE}
# selected_combined <- read_rds(file="../Results/randomforestSRC/selectedCombined_top7_baseline_adult.rds")
selected_cytokines <- c("IL.17A")
selected_metabols <- c("TYROSINE" ,"MALATE" ,"EICOSAPENTAENOIC ACID" ,"ARACHIDONIC ACID" ,"VALINE")


data <- as.data.frame(cbind(cytokineData.train[,selected_cytokines,drop=FALSE],metaboData.train[,selected_metabols,drop=FALSE]))
data$Class = as.factor(gsub(" ","",sampleData.train$Outcome))
data$Class <- factor(data$Class, levels =c("CURED","FAILED"))

data_test <- as.data.frame(cbind(cytokineData.test[,selected_cytokines,drop=FALSE],metaboData.test[,selected_metabols,drop=FALSE]))
data_test$Class = as.factor(gsub(" ","",sampleData.test$Outcome))
data_test$Class <- factor(data_test$Class, levels =c("CURED","FAILED"))

data_test_HIVpos <- as.data.frame(cbind(cytokineData.test.HIVpos[,selected_cytokines,drop=FALSE],metaboData.test.HIVpos[,selected_metabols,drop=FALSE]))
data_test_HIVpos$Class = as.factor(gsub(" ","",sampleData.test.HIVpos$Outcome))
data_test_HIVpos$Class <- factor(data_test_HIVpos$Class, levels =c("CURED","FAILED"))

data_test_HIVneg <- as.data.frame(cbind(cytokineData.test.HIVneg[,selected_cytokines,drop=FALSE],metaboData.test.HIVneg[,selected_metabols,drop=FALSE]))
data_test_HIVneg$Class = as.factor(gsub(" ","",sampleData.test.HIVneg$Outcome))
data_test_HIVneg$Class <- factor(data_test_HIVneg$Class, levels =c("CURED","FAILED"))
```

#### Train SVM model

```{r}
set.seed(123)
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)  

svmGrid <- expand.grid(sigma = seq(0.01, 1, length.out = 10),
                       C = 2^seq(-5, 10, length.out = 10))

cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)

svmTune <- train(Class ~ ., data = data, 
                 method = "svmRadial",  
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 tuneGrid = svmGrid,
                 trControl = fitControl)

stopCluster(cl)
```

#### Test and ROC {.tabset}
##### Train data
###### Test SVM 

```{r}
pred_combined_train <- predict(svmTune, data)
prob_combined_train <- predict(svmTune, data, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data$Class, 
            predictor = prob_combined_train[,1], 
            levels = rev(levels(data$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_train <- roc0
```


##### Entire test data
###### Test SVM 

```{r}
pred_combined_test <- predict(svmTune, data_test)
prob_combined_test <- predict(svmTune, data_test, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test$Class, 
            predictor = prob_combined_test[,1], 
            levels = rev(levels(data_test$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])


roc_combined_test <- roc0
```



##### Entire test data (only HIV pos)
###### Test SVM svmTune 

```{r}
pred_combined_test_HIVpos <- predict(svmTune, data_test_HIVpos)
prob_combined_test_HIVpos <- predict(svmTune, data_test_HIVpos, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVpos$Class, 
            predictor = prob_combined_test_HIVpos[,1], 
            levels = rev(levels(data_test_HIVpos$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_HIVpos <- roc0
```










##### Entire test data (only HIV neg)
###### Test SVM svmTune 

```{r}
pred_combined_test_HIVneg <- predict(svmTune, data_test_HIVneg)
prob_combined_test_HIVneg <- predict(svmTune, data_test_HIVneg, type="prob")
```

###### ROC curve

```{r}
roc0 <- roc(response = data_test_HIVneg$Class, 
            predictor = prob_combined_test_HIVneg[,1], 
            levels = rev(levels(data_test_HIVneg$Class)),ci=TRUE)

ciobj <- ci.se(roc0, specificities=seq(0, 1, l=25))
dat.ci <- data.frame(x = as.numeric(rownames(ciobj)),
                     lower = ciobj[, 1],
                     upper = ciobj[, 3])

roc_combined_test_HIVneg <- roc0
```










##### Both train and test

```{r}
bothrocPlot_func(roc_combined_train,roc_combined_test)
```

##### Both train and test and both HIV pos and HIV neg

```{r}
fourrocPlot_func(roc_combined_train,roc_combined_test,roc_combined_test_HIVpos,roc_combined_test_HIVneg)
result <- threerocPlot_func_HIV(roc_combined_test,roc_combined_test_HIVpos,roc_combined_test_HIVneg)
result$plot
```

### Genarate figures
```{r}
result <- threerocPlot_func(roc_combined_test, roc_cytokine_test,roc_metabol_test,
                            names=c("Combined","Cytokine","Metabolite"),colors=c("#50b957", "#d84882","#777acd"))
result$plot  
Fig4_total <- result$plot 
result <- threerocPlot_func(roc_combined_test, roc_combined_test_HIVpos, roc_combined_test_HIVneg, names=c("Combined","HIV-positive","HIV-negative"),colors=c("#50b957", "red","#d89f32"))
result$plot  
Fig4_combined_HIV <- result$plot 

result <- threerocPlot_func(roc_cytokine_test, roc_cytokine_test_HIVpos, roc_cytokine_test_HIVneg,names=c("Cytokine","HIV-positive","HIV-negative"),colors=c("#d84882",  "red","#d89f32"))
result$plot  
Fig4_cytokine_HIV <- result$plot 

result <- threerocPlot_func(roc_metabol_test, roc_metabol_test_HIVpos, roc_metabol_test_HIVneg,
                            names=c("Metabolite","HIV-positive","HIV-negative"),colors=c("#7b63d4", "red","#d89f32"))
result$plot  
Fig4_metabol_HIV <- result$plot 

save(Fig4_metabol_HIV,Fig4_cytokine_HIV,Fig4_total, Fig4_combined_HIV,file="../Results/Fig4-ROCs-treatment.RData")
```

## Data at baseline (only adults) - Enrichment of 6 markers treatment response
### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult_TreatmentFollowUp.RData")
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
MetaboData_Baseline_Adult_TreatmentFollowUp <- log2(MetaboData_Baseline_Adult_TreatmentFollowUp+1)%>%as.data.frame()
selected_cytokines <-c("IL.17A")
selected_metabols <- c("TYROSINE","ARACHIDONIC ACID", "MALATE" ,"EICOSAPENTAENOIC ACID" ,"VALINE")

SA_samples <- read_excel("../../Data/sa samples_JMC.xlsx")
SA_samples$LONGID <- gsub("SA","",SA_samples$ID)
ClinicalData_Baseline_Adult[match(SA_samples$ID,ClinicalData_Baseline_Adult$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult%>%filter(Outcome!="EXCLUDE")

ClinicalData_Baseline_Adult_TreatmentFollowUp[match(SA_samples$ID,ClinicalData_Baseline_Adult_TreatmentFollowUp$ID),"Outcome"] <- SA_samples$`New Group`
ClinicalData_Baseline_Adult_TreatmentFollowUp <- ClinicalData_Baseline_Adult_TreatmentFollowUp%>%filter(Outcome!="EXCLUDE")
ClinicalData_Baseline_Adult_TreatmentFollowUp[which(ClinicalData_Baseline_Adult_TreatmentFollowUp$Visit.Week==17|ClinicalData_Baseline_Adult_TreatmentFollowUp$Visit.Week==26),"Visit.Week"] <- "17/26"

inputData <- CytokineData_Baseline_Adult_TreatmentFollowUp_imputed[ClinicalData_Baseline_Adult_TreatmentFollowUp$ID,selected_cytokines,drop=FALSE]
inputData$Visit.Week <- factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Visit.Week, levels=c("0","8","17/26"))
inputData$Outcome <- as.character(factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Outcome, levels=c("CURED","FAILED")))
inputData$LONGID <- as.character(ClinicalData_Baseline_Adult_TreatmentFollowUp$LONGID)
inputData_melted <- melt(inputData)
inputData_melted$Type="Cytokine"

inputData <- MetaboData_Baseline_Adult_TreatmentFollowUp[ClinicalData_Baseline_Adult_TreatmentFollowUp$ID,selected_metabols]
inputData$Visit.Week <- factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Visit.Week, levels=c("0","8","17/26"))
inputData$Outcome <- as.character(factor(ClinicalData_Baseline_Adult_TreatmentFollowUp$Outcome, levels=c("CURED","FAILED")))
inputData$LONGID <- as.character(ClinicalData_Baseline_Adult_TreatmentFollowUp$LONGID)
inputData_melted_temp <- melt(inputData)
inputData_melted_temp$Type="Metabol"
inputData_melted_temp$variable <- tolower(inputData_melted_temp$variable)
inputData_melted <- rbind(inputData_melted,inputData_melted_temp)
inputData_melted <- inputData_melted%>%filter(!is.na(Outcome))


ClinicalData_Baseline_Adult_noTB <- ClinicalData_Baseline_Adult%>%filter(TB=="No")
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
selected_cytokines <-c("IL.17A")
selected_metabols <- c("TYROSINE","ARACHIDONIC ACID", "MALATE" ,"EICOSAPENTAENOIC ACID" ,"VALINE")
inputData <- CytokineData_Baseline_Adult_imputed[ClinicalData_Baseline_Adult_noTB$ID,selected_cytokines,drop=FALSE]
inputData$Visit.Week <- "ORD"
inputData$Outcome <- "ORD"
inputData$LONGID <- NA
inputData_melted_temp <- melt(inputData)
inputData_melted_temp$Type="Cytokine"

inputData <- MetaboData_Baseline_Adult[ClinicalData_Baseline_Adult_noTB$ID,selected_metabols]
inputData$Visit.Week <- "ORD"
inputData$Outcome <- "ORD"
inputData$LONGID <- NA
inputData_melted_temp2 <- melt(inputData)
inputData_melted_temp2$Type="Metabol"
inputData_melted_temp2$variable <- tolower(inputData_melted_temp2$variable)
inputData_melted <- rbind(inputData_melted,inputData_melted_temp,inputData_melted_temp2)
```

### Visualization

```{r}
outcome_colors <- c("ORD" = "#ca6332","CURED" = "#3d814b", "FAILED" = "#cf3f42")  

plot <- ggplot(inputData_melted, aes(x = factor(Visit.Week), y = value)) +
  geom_boxplot(outlier.shape = NA, color = "black", fill = "snow") +
  geom_point(aes(color = Outcome), size = 2, position = position_jitter(width = 0.2)) +
  scale_color_manual(values=outcome_colors,labels=c("ORD"="ORD","CURED"="PTB-CURED","FAILED"="PTB-FAILED"))+
  facet_wrap(~ variable, scales = "free_y") +
  theme_classic(base_size = 26) +
  theme(
    legend.text = element_text(size = 26), 
    axis.title.x = element_text(size = 26, face="bold"),
    axis.title.y = element_text(size = 26, face="bold"),
    strip.text = element_text(face = "bold", size = 26),
    legend.position = "top"
  ) +
  labs(x = "Visit Week", y = expression(bold(log[2](Plasma_Concentration))), color = "")

pdf(file="../Results/Figures/Manuscript/Figure 4-b-1.pdf",width = 21,height = 10)
plot
dev.off()

plot <- ggplot(inputData_melted, aes(x = factor(Visit.Week), y = value, color = Outcome)) +
  geom_boxplot(aes(fill = Outcome), 
               outlier.shape = NA, 
               position = position_dodge(width = 0.8),
               alpha=0.3) +
  geom_point(size = 2, 
             position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8)) +
  scale_color_manual(values = outcome_colors,
                     labels = c("ORD"="ORD","CURED"="PTB-CURED","FAILED"="PTB-FAILED")) +
  scale_fill_manual(values = outcome_colors,
                    labels = c("ORD"="ORD","CURED"="PTB-CURED","FAILED"="PTB-FAILED")) +
  facet_wrap(~ variable, scales = "free_y") +
  theme_classic(base_size = 26) +
  theme(
    legend.text = element_text(size = 26), 
    axis.title.x = element_text(size = 26, face="bold"),
    axis.title.y = element_text(size = 26, face="bold"),
    strip.text = element_text(face = "bold", size = 26),
    legend.position = "top"
  ) +
  labs(x = "Visit Week", y = expression(bold(log[2](Plasma_Concentration))), color = "", fill = "")


pdf(file="../Results/Figures/Manuscript/Figure 4-b-2.pdf",width = 21,height = 10)
plot
dev.off()
```

## Make Figure 4

```{r}
load("../Results/Fig4-ROCs-treatment.RData")
pdf(file="../Results/Figures/Manuscript/Figure 4.pdf",width=32,height = 20)
plot_grid(Fig4_total,NULL,NULL,ncol=3,labels = c("a"),label_size = 36)/plot_grid(Fig4_combined_HIV,Fig4_cytokine_HIV,Fig4_metabol_HIV,ncol = 3,labels=c("c","d","e"),label_size = 36)
dev.off()
```


## Data at baseline (only adults) - Pathway enrichment 

### Ion mode {.tabset}

#### Positive mode 

##### Read feature table 

```{r}
feature_table <- read.delim("../../Data/Quantified Metabolomics Data/Untargeted/ComBat_mzcalibrated_untargeted_mediansummarized_featuretable_hilicpos.txt")
l <- load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")

feature_table <- feature_table[,c(1,2,11:ncol(feature_table))]
feature_table%>%colnames()
feature_table <- feature_table[,c(1,2,which(colnames(feature_table)%in%ClinicalData_Baseline_Adult$hilicposFileName))]
which(!feature_table%>%colnames()%in%ClinicalData_Baseline_Adult$hilicposFileName)
feature_table <- feature_table[,c("mz","time",ClinicalData_Baseline_Adult$hilicposFileName)]
any(is.na(feature_table))
```

##### Run linear model on each feature using p-values

```{r}
feature_table_value <- feature_table[,-c(1,2)]  # numeric features

covareiateData <- ClinicalData_Baseline_Adult[,c("AGE","HIV","Sex"),drop=FALSE]
f <- factor(ClinicalData_Baseline_Adult$TB)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("noTB","TB","Age","HIV","Sex")
cont.matrix <- makeContrasts(TB-noTB,levels=design)
fit <- lmFit(feature_table_value, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)


ttest_df <- as.data.frame(cbind(fit.b$p.value,fit.b$t))

ttest_df$adj_p <- p.adjust(fit.b$p.value, method = "BH")
ttest_df <- cbind(feature_table[,c("mz","time")],ttest_df)
head(ttest_df)
names(ttest_df) <- c("m/z","retention time","p-value","statistic","adj-p")
ttest_df_input <- ttest_df%>%select(`m/z`,`retention time`,`p-value`,statistic)
ttest_df_input <- ttest_df_input%>%arrange(`p-value`)
write.table(ttest_df_input,file="../Results/Mummichog/lineartest-pos.txt",sep = "\t",quote = FALSE,row.names=FALSE)
```


##### Visualization with Manhattan plot

```{r}
Fig5_pos_manhat_1 <- manhattan_plot(ttest_df_input,alpha_threshold = 0.0001,x_axis = "m/z")
Fig5_pos_manhat_2 <- manhattan_plot(ttest_df_input,alpha_threshold = 0.0001,x_axis = "retention time")
```
#### Negative mode

##### Read feature table 

```{r}
feature_table <- read.delim("../../Data/Quantified Metabolomics Data/Untargeted/ComBat_mzcalibrated_untargeted_mediansummarized_featuretable_c18neg.txt")
l <- load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")

feature_table <- feature_table[,c(1,2,11:ncol(feature_table))]
feature_table%>%colnames()
feature_table <- feature_table[,c(1,2,which(colnames(feature_table)%in%ClinicalData_Baseline_Adult$c18negFileName))]
which(!feature_table%>%colnames()%in%ClinicalData_Baseline_Adult$c18negFileName)
feature_table <- feature_table[,c("mz","time",ClinicalData_Baseline_Adult$c18negFileName)]
any(is.na(feature_table))
```

##### Run linear model on each feature using p-values

```{r}
feature_table_value <- feature_table[,-c(1,2)]  # numeric features

covareiateData <- ClinicalData_Baseline_Adult[,c("AGE","HIV","Sex"),drop=FALSE]
f <- factor(ClinicalData_Baseline_Adult$TB)
design <- model.matrix(~0+f+AGE+HIV+Sex,data=covareiateData)
colnames(design) <- c("noTB","TB","Age","HIV","Sex")
cont.matrix <- makeContrasts(TB-noTB,levels=design)
fit <- lmFit(feature_table_value, design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit.b <- eBayes(fit2)


ttest_df <- as.data.frame(cbind(fit.b$p.value,fit.b$t))

ttest_df$adj_p <- p.adjust(fit.b$p.value, method = "BH")
ttest_df <- cbind(feature_table[,c("mz","time")],ttest_df)
head(ttest_df)
names(ttest_df) <- c("m/z","retention time","p-value","statistic","adj-p")
ttest_df_input <- ttest_df%>%select(`m/z`,`retention time`,`p-value`,statistic)
ttest_df_input <- ttest_df_input%>%arrange(`p-value`)
write.table(ttest_df_input,file="../Results/Mummichog/lineartest-neg.txt",sep = "\t",quote = FALSE,row.names=FALSE)
```


##### Visualization with Manhattan plot

```{r}
Fig5_neg_manhat_1 <- manhattan_plot(ttest_df_input,alpha_threshold = 0.0001,x_axis = "m/z")
Fig5_neg_manhat_2 <- manhattan_plot(ttest_df_input,alpha_threshold = 0.0001,x_axis = "retention time")
```

### Pathways enriched using linearmodel results

```{r}
Mummichog_pos <- read_excel("../Results/Mummichog/lineartest-pos/tables/mcg_pathwayanalysis_lineartest-pos.xlsx")
Mummichog_neg <- read_excel("../Results/Mummichog/lineartest-neg/tables/mcg_pathwayanalysis_lineartest-neg.xlsx")
Mummichog_pos$`adjp-value` <- p.adjust(Mummichog_pos$`p-value`,method = "BH")
Mummichog_neg$`adjp-value` <- p.adjust(Mummichog_neg$`p-value`,method = "BH")
Mummichog_sig <- Mummichog_pos%>%filter(`p-value`<0.05)%>%mutate(ion="pos")%>%
  bind_rows(Mummichog_neg%>%filter(`p-value`<0.05)%>%mutate(ion="neg"))
Mummichog_sig <- Mummichog_sig%>%filter(`adjp-value`<= 0.05)%>%mutate(highlight="Yes")%>%
  bind_rows(Mummichog_sig%>%filter(`adjp-value`> 0.05)%>%mutate(highlight="No"))

Mummichog_sig$logP <- -log10(Mummichog_sig$`p-value`)
Mummichog_sig <- Mummichog_sig%>%arrange(`p-value`)
Mummichog_sig$pathway <- factor(Mummichog_sig$pathway,levels=unique(Mummichog_sig$pathway))

Mummichog_sig <- Mummichog_sig %>%
  mutate(label_color = ifelse(highlight == "Yes", "red", "black"))

Fig5_pathways <- ggplot(Mummichog_sig, aes(y = pathway, x = logP)) +
  geom_segment(aes(xend = 0, yend = pathway), color = "grey70", linewidth = 0.6) +
  geom_point(aes(color = ion), size = 12, alpha = 0.9) +
  geom_text(aes(label = overlap_size), size = 12,  color="black",face="bold",vjust = 0.5, hjust = 0.5) +
  geom_vline(xintercept = -log10(0.05), color = "red", linetype = "dashed", linewidth = 1) +
  scale_color_manual(values = c("#bf4eb4", "#89b436"), labels=c("pos"="Positive","neg"="Negative"),name = "Ion Mode: ") +
  scale_x_continuous(expand = c(0.01, 0), limits = c(0, 4.5)) +
  labs(x = expression(bold(-log[10](FDR-corrected_p-value))), y = NULL) +
  theme_classic(base_size = 26) +
  theme(
    axis.text.y = element_text(size = 26,color=Mummichog_sig$label_color),
    axis.text.x = element_text(size = 26),
    axis.title.x = element_text(size = 26, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 26)
  )
pdf(file="../Resfults/Figures/Manuscript/Figure 5-c-1.pdf",width = 15,height = 10)
Fig5_pathways
dev.off()

pos_list <- Mummichog_sig%>%filter(ion=="pos")%>%select(pathway)%>%unlist()
neg_list <- Mummichog_sig%>%filter(ion=="neg")%>%select(pathway)%>%unlist()

venn_data <- list(
  Negative = neg_list,
  Positive = pos_list
)

library(ggvenn)
vennPlot <- ggvenn(
  venn_data,
  fill_color = c("#bf4eb4", "#89b436"),
  set_name_size = 26,  
  text_size = 26      
) 
pdf(file="../Results/Figures/Manuscript/Figure 5-c-2.pdf",width = 20,height = 20)
vennPlot
dev.off()
```

## Data at baseline (only adults) - correlation plot

### Load data

```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
load("../Results/DEA/DEC_Baseline_Adult.RData")
load("../Results/DEA/DEM_Baseline_Adult.RData")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult[match(row.names(CytokineData_Baseline_Adult),ClinicalData_Baseline_Adult$ID),]
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult <- MetaboData_Baseline_Adult[ClinicalData_Baseline_Adult$ID,]

input_metabolites <- volcanoPlot_data_MET%>%filter(qValue.TB.noTB<0.05)%>%select(Metabolite)%>%unlist()
input_cytokines <- volcanoPlot_data_CYT%>%filter(qValue.TB.noTB<0.05)%>%select(Cytokine)%>%unlist()
input_metaboData <- MetaboData_Baseline_Adult[,input_metabolites]
input_cytoData <- CytokineData_Baseline_Adult_imputed[,input_cytokines]
```

### Visualization using Chord plot

```{r}
res <- rcorr(as.matrix(input_metaboData), as.matrix(input_cytoData), type = "pearson")

cor_mat <- res$r
p_mat   <- res$P

metabolite_names <- colnames(input_metaboData)
cytokine_names   <- colnames(input_cytoData)

cor_sub <- cor_mat[metabolite_names, cytokine_names]
p_sub   <- p_mat[metabolite_names, cytokine_names]

df_long <- melt(cor_sub, varnames = c("Metabolite", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_sub)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_sig <- df_long%>%filter(adj.p.value<0.05, abs(Correlation)>0.3)

links <- data.frame(from = df_sig$Metabolite,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

col_fun <- colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))

grid.col <- c(setNames(rep("#7b63d4", length(metabolite_names)), tolower(metabolite_names)),
              setNames(rep("#d84882", length(cytokine_names)), cytokine_names))

library(circlize)

links$from <- tolower(links$from)

pdf("../Results/Figures/Manuscript/Figure5-d.pdf", width = 14, height = 10)

circos.par(
  track.margin = c(0.01, 0.01),
  canvas.xlim = c(-1.4, 1.4),
  canvas.ylim = c(-1.4, 1.4)
)

chordDiagram(
  links,
  transparency = 0.4,
  grid.col = grid.col,
  col = col_fun(links$value),
  annotationTrack = "grid",
  preAllocateTracks = list(track.height = 0.05)
)

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name = get.cell.meta.data("sector.index")
    
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 2,
      font = 1
    )
  },
  bg.border = NA
)

circos.clear()
dev.off()
```




## Data at baseline (only adults) - box plot for fatty acid markers
### Load data
```{r}
load("../../Data/ReadyData/ReadyData_Baseline_Adult.RData")
ClinicalData_Baseline_Adult <- ClinicalData_Baseline_Adult[match(row.names(CytokineData_Baseline_Adult),ClinicalData_Baseline_Adult$ID),]
MetaboData_Baseline_Adult <- log2(MetaboData_Baseline_Adult+1)%>%as.data.frame()
MetaboData_Baseline_Adult <- MetaboData_Baseline_Adult[ClinicalData_Baseline_Adult$ID,]

fatty_acids <- c(
  "ARACHIDIC ACID",
  "ARACHIDONIC ACID",
  "DOCOSAHEXAENOIC ACID",
  "DOCOSATETRAENOIC ACID",
  "EICOSADIENOIC ACID",
  "EICOSAPENTAENOIC ACID",
  "GONDOIC ACID",
  "HOMOLINOLENIC ACID",
  "LINOLENIC ACID",
  "PALMITOLEIC ACID"
)
input_data <- MetaboData_Baseline_Adult[,fatty_acids]
input_data$TB <- ClinicalData_Baseline_Adult$TB
input_data <- melt(input_data)
input_data$variable <- tolower(input_data$variable)
boxplot <- boxplot_func(input_data,xaxis = "variable",yaxis = "value",group = "TB"  )
pdf(file="../Results/Figures/Manuscript/Figure 5-e.pdf",width = 20, height = 15)
boxplot
dev.off()

```

## Make Figure 5

```{r}
pdf(file="../Results/Figures/Manuscript/Figure 5.pdf",height = 10,width = 32)
plot_grid(Fig5_neg_manhat_1,Fig5_pos_manhat_1,Fig5_pathways, ncol = 3, rel_widths = c(0.5,0.5,1.75))
dev.off()
```





